### 2022-07-07 GroupPolarizationStatmod/papers/ms.tex
3/10 analyzed; saving all data, means displayed correctly

% Author: Matthew Turner

% \documentclass[11pt,letterpaper]{scrartcl}
\documentclass[letterpaper,man,natbib]{apa6}
% \documentclass[11pt]{report}
% \documentclass{report}
% \documentclass{book}
\usepackage[bookmarks, hidelinks]{hyperref}
\usepackage{amssymb,amsmath}
% \usepackage{fullpage}
\usepackage{tabulary}
\usepackage{tabularx}
\usepackage{float}
% \usepackage[margin=1.00in]{geometry}
% \usepackage[margin=0.90in]{geometry}

\usepackage{caption}
\usepackage{booktabs}
\usepackage{pslatex}
\usepackage{apacite}
\usepackage{authblk}
\usepackage{subcaption}
\usepackage{pgfplots}
\usepackage{wrapfig}
\usepackage[english]{babel}
\usepackage{lmodern}
% \usepackage{setspace}
% \doublespace
% \usepackage{url}
\usepackage{bigfoot}
\usepackage[export]{adjustbox}
% \setlength\intextsep{0pt}

\usepackage{graphicx}

\title{How simple consensus to the mean can be mistaken for ``group polarization'' extremity-shifted consensus}
\shorttitle{Group polarization or simple consensus?}
\author[1]{{Matthew A.~Turner}}
\affil[1]{Department of Earth System Science, Stanford University}

\author[2,3]{{Paul E.~Smaldino}}
\affil[2]{Cognitive and Information Sciences, University of California, Merced}
\affil[3]{Center for Advanced Study in the Behavioral Sciences, Stanford University}

\abstract{
  Group polarization is the name given by social scientists for the observed tendency of
  like-minded groups to become more extreme as they discuss some topic of interest.
  Clearly it is important to understand such a phenomenon to understand broader
  problematic trends towards increased polariztion across society as a whole.
  This group-level shift in opinion extremity is often measured 
  Unfortunately, recent analyses have demonstrated that this measurement and
  analysis approach can result in false positive detections of changes in mean
  that are actually due to floor/ceiling effects from using ordinal scales to
  measuring continuously-distributed latent psychological variables, and not
  accounting for this mismatch. In this paper we analyze a selection of influential
  group polarization studies to demonstrate that many group polarization 
  detections reported throughout decades of literature are plausibly false
  due to these floor/ceiling effects. We develop a generative model of group
  polarization data, use it to represent each group polarization study, and
  investigate whether we can generate observed opinion shifts to extremity by
  setting the latent opinion mean to be constant, but allow latent opinion variance
  to decrease from pre- to post-discussion, which theoritcally 
  occurs in \emph{simple consensus} where the mean stays constant. After we identify which
  positive detections of group polarization are plausibly false in this way, we further
  investigate whether the latent constant mean opinion and pre- and post-discussion
  opinion variances can fool the statistical tests used in each study to 
  generate false positives. Finally, we demonstrate how a more sophisticated
  thresholded-normal statistical model can account for the metric-ordinal 
  mismatch to estimate latent distribution parameters and accurately identify
  simple consensus.
}

\begin{document}
\maketitle
\setlength{\parindent}{15pt}
% \maketitle

\section{Introduction}

\subsection{Group polarization}

\subsection{Statistical inference in opinion change}

2-3 par: the right way; what GP studies do; why it's a problem (N-1 sents) and how
to fix it (1 sent)

When

\begin{itemize}
  \item 
    Group polarization studies use null-hypothesis testing that implicitly
    assume (1) \emph{opinions} (or other equivalent latent psychological variables
    such as \emph{attitudes} or \emph{beliefs}) are continuously-valued, and
    (2) 
  \item
    The exact null-hypothesis testing procedure varies between studies $t$-tests, ANOVA, etc. 
\end{itemize}

Recently, problems with using have been made clear~\citep{Liddell2018}.
The way to fix this problem going forward is to use a thresholded-normal 
statistical model for group polarization data, or other related statistical 
model that accounts for opinion binning~\citep[Ch. 23]{Liddell2018,KruschkeDBDA}.

\subsection{Study selection and overview}

Studies were selected to be either influential papers, based on citations or
author prominence, that test either examine
group polarization in political opinions, and to be representative of investigations
of four primary group polarization explanations extant in the literature.
At the time of paper selection we did not examine the statistical methods in detail beyond
checking that they indeed used

Because of the difficulties casting individual research articles into our modeling
framework, and the central task of creating the model and analyzing this first
batch of studies with it, we limit our analysis to 10 studies. It is problem
enough that these 10 prominent studies have the problems we demonstrate they
plausibly do. Problems among these 10 suggest the problem may be more wide spread. 
When we do more analyses
they will be aided by improved data management systems for applying this model to 
additional studies (eventually outside of group polarization).

Statistical problems among these studies are not limited to a mismatch between
measurement and statistical procedures, which requires the detailed treatment
we provide in this paper. Other problems are more pedestrian.
One major issue shared by several studies
is the use of unpaired $t$-tests to detect group polarization. Since the same
group has their opinions measured twice over time, the opinions are correlated
and so a paired $t$-test would be required for test assumptions to be met 
(assuming all other paired $t$-test assumptions were met, 
which of course they are not, which is why our paper here is necessary).
In some cases the reported shifts were ``significant'', 
but not truly ``group polarization'' shifts, i.e., 
shifts from initially biased group opinions
to more extreme group opinions biased in the same direction; instead shifts 
were from a neutral initial group opinion to a biased group opinion 
(as in, e.g.,~\citet{Abrams1990} Uncategorized2), or from
an opinion biased in one direction (as in, e.g.,~\citet{Abrams1990} Categorized1,
Categorized2, AND OTHERS???). In some results of \citet{Freidkin1999a}, 
shift values are reported with their
standard deviations---but the shift standard deviations are greater than the
shifts themselves, indiciating that zero shift is included in the observed
shift confidence interval. This is \emph{prima facie} a negative result that
should not have been counted as evidence of group polarization.

Some group polarization studies do not measure continuous latent opinions using
a metric scale, and so are not affected by the floor and ceiling effects that
undermine a good number of experimental results, as we examine here. There may,
however, be other problems, such as non-group polarization shifts and the use
of unpaired $t$-tests when paired tests are required.

\subsection{Research overview}

In this paper we analyze whether published statistical detections of group 
polarization obtained through null-hypothesis statistical tests
are plausibly false because they can be explained equally well by simple consensus
where the mean group opinion does not change over time, but opinion variance does.
This analysis requires several preliminary steps. 
First we formally develop our generative model of simple consensus and group 
polarization in terms of changes in group opinion mean and variance from 
pre- to post-deliberation. Then we developed a search algorithm for identifying 
which, if any, simple consensus scenarios with 
constant latent mean and pre- and post-deliberation opinion variances 
could generate published observations of group polarization. To apply this model
and algorithm to published results, we developed an interactive web application
for data input and to test which different potential simple consensus scenarios 
could generate group polarization opinion shifts. After identifying plausibly false
detections of group polarization we examine whether the identified simple
consensus scenario (parameters?) can indeed generate data to fool the 
paper's null-hypothesis tests into thinking the data actually support the inference
that simple consensus did not occur. Finally, we examine whether, in these same
plausibly false detections, a Bayesian thresholded-normal statistical model
will accurately identify simple consensus instead of falsely detecting group
polarization.

\section{Model}

Our model represnts the generative process of finding different forms of group consensus
in group polarization experiments, and the process of measuring that consensus using
an ordinal measurement scale, e.g.\ a Likert-type scale. This model considers two
relevant forms of consensus: \emph{simple consensus} and \emph{group polarization}.
Either process is \emph{consensus} since opinion variance among the group 
decreases as individuals come to agree more than they did previously. 
Group polarization experiments are designed to foster group consensus
it is not necessary to consider clustering, negative influence, or other
processes known to generally occur when opinions within a group significantly
diverge~\citep{Turner2018}.

\section{Analysis}

\bibliographystyle{apacite}

\setlength{\bibleftmargin}{.125in}
\setlength{\bibindent}{-\bibleftmargin}

\bibliography{/Users/mt/workspace/Writing/library.bib}

\appendix

\section{Notes on studies}

\paragraph{\citet{Abrams1990}}

\paragraph{\citet{Burnstein1973}}

\paragraph{\citet{Hogg1990}} It is not clear if there are indeed significant pre-post
deliberation differences reported in this paper. The Results are a tangle of
estimates of effect sizes and diagnostic statistical measures. It is also
not clear that they actually report the pre- and post-deliberation opinions;
it appears they only report the shifts. ``Opinion Difference Measures'' 
does not seem to include a test that the shifts themselves are reliable.

\end{document}