### 2024-04-09 GroupPolarizationStatmod/paper.bk/tables/conditions_table.tex
paper rewrite

% \newgeometry{margin=1cm} % modify this if you need even more space
\begin{landscape}

%put your table here

% \begin{table}
%   \small
%   \begin{tabular}{p{1.5in}lp{2.5in}p{3.5in}}
  % \begin{longtable}{p
  \small
  \begin{longtable}{p{1.5in}lp{2.0in}p{4.0in}}
    \caption{Summary of idiosyncracies across experimental conditions in the
    studies we scrutinized for plausibly-false positives.}
    \label{tab:conditionsComparison}\\
  % \begin{longtable}{llll}
  % \begin{tabular}{llll}
  \toprule
  \textbf{Article (Article Tag)} & \textbf{Condition Tag} & \textbf{Prompt} & \textbf{Description} \\
  \midrule
  \citeA{Moscovici1969} (Moscovici1969) & Americans-Opinion & American economic aid is always used for
  political pressure.  & 11 total prompts (1 reported), all 7-point scales. Unclear how data was aggregated. \\
  \midrule
  ~(Moscovici1969) & DeGaulle-Opinion & Charles De Gaulle is too old to (be president) & 12 total prompts (1 reported), all 7-point scales. Unclear how data was aggregated. \\
  \midrule
  ~(Moscovici1969) & DeGaulle-Judgment & Charles De Gaulle is too old to (be president) & Same as
  above, except ``subjects were asked first to judge individually and then to reach consensus as a
  group in evaluating the varoableness or unfavorableness of each item toward de Gaulle,
  regardless of whether they agreed...~with the content of the item.'' \\
  \bottomrule
  \end{longtable}
  % \end{tabular}
% \end{table}

\end{landscape}
\restoregeometry