### 2024-04-09 GroupPolarizationStatmod/paper/_extensions/partials/title.tex
init quarto paper rewrite

\usepackage{authblk}
\usepackage{etoolbox}
\makeatletter
\title{$title$}
\makeatother
\author[1,*]{{Matthew A.~Turner}}
\affil[1]{\small Division of Social Sciences, Stanford Doerr School of Sustainability, Stanford University}

\author[2,3]{{Paul E.~Smaldino}}
\affil[2]{\small Cognitive and Information Sciences, University of California, Merced}
\affil[3]{\small Santa Fe Institute} 
% \vspace{2em}
\affil[*]{\small Correspondence: \href{mailto:maturner@stanford.edu}{maturner@stanford.edu}}

\date{}