### 2024-04-11 GroupPolarizationStatmod/paper/ms.qmd
Updated abstract and title, stub intro

---
title: "Foundational claims of group polarization are plausibly false, with lessons for studying psychological dynamics"
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    cite-method: biblatex
    biblatexoptions: sorting=none, style=numeric-comp
    pdf-engine: latexmk
    pdf-engine-opt: -xelatex
    keep-tex: true
    template-partials:
      - _extensions/partials/title.tex
authors:
  - name: Matthew Adam Turner
    affiliation: Stanford University Doerr School of Sustainability, Stanford University
    roles: writing, model development, project development
    corresponding: true
    email: maturner@stanford.edu
  - name: Paul E. Smaldino
    affiliation: 
      - Cognitive and Information Sciences, Univeristy of California, Merced
      - Santa Fe Institute
    roles: writing
    corresponding: true
bibliography: ../library.bib
classoption: abstract
keep-tex: true
abstract: |
  \noindent _Group polarization_ is the accepted name for a form of consensus where members of a like-minded group become, on average, more extreme in their opinions after discussing a topic. Group polarization is important because it may drive broader polarization if relatively moderate groups become more extreme. Group polarization is celebrated as a regular, reliable outcome of a certain experimental design based on dozens of published studies that purportedly demonstrate the phenomena. However, in this paper we show that 55 of 60 published observations of purported group polarization, from ten foundational studies, are equally well-explained by a simple consensus model where group opinion variance decreases, but extremity does not increase. We show this through a novel computational meta-analysis framework designed to identify plausible pre- and post-discussion opinion distributions that are consistent with simple consensus, but appear to be group polarization when measured on ordinal, Likert-style scales. This occurs due to ceiling effects induced by ordinal measurements. At best this shows that a selection of foundational findings in social psychology must be discarded. At worst this is the tip of the iceberg, and many similar social psychological findings within and without the group polarization tradition will similarly be shown to be unreliable.
geometry:
  - margin=1.1in
  - letterpaper
biblio-title: References
---

## Introduction