### 2024-04-16 GroupPolarizationStatmod/paper/ms.qmd
Intro outlined; nearly complete

---
title: "Foundational claims of group polarization are plausibly false, with lessons for the empirical study of psychological dynamics"
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    cite-method: biblatex
    biblatexoptions: sorting=none, style=numeric-comp
    pdf-engine: latexmk
    pdf-engine-opt: -xelatex
    keep-tex: true
    template-partials:
      - _extensions/partials/title.tex
authors:
  - name: Matthew Adam Turner
    affiliation: Stanford University Doerr School of Sustainability, Stanford University
    roles: writing, model development, project development
    corresponding: true
    email: maturner@stanford.edu
  - name: Paul E. Smaldino
    affiliation: 
      - Cognitive and Information Sciences, Univeristy of California, Merced
      - Santa Fe Institute
    roles: writing
    corresponding: true
bibliography: "/Users/mt/workspace/Writing/library.bib"
classoption: abstract
keep-tex: true
geometry:
  - margin=1.1in
  - letterpaper
biblio-title: References
indent: true
abstract: |
  \noindent _Group polarization_ is the name for a form of consensus where members of a like-minded group become, on average, more extreme in their opinions after discussing a topic. Group polarization is important because it may drive political polarization if moderate partisans become more extreme. Theoretically, group polarization is celebrated as a regular, reliable outcome of a certain experimental design based on dozens of published studies that purportedly demonstrate the phenomena. However, in this paper we show that 55 of 60 published observations of purported group polarization, from ten foundational studies, are equally well-explained by a simple consensus model where group opinion variance decreases, but extremity does not increase. We use a novel computational meta-analysis framework designed to identify plausible pre- and post-discussion opinion distributions that are consistent with simple consensus, but appear to be group polarization when measured on ordinal, Likert-style scales. This occurs due to ceiling effects induced by ordinal measurements. At best this shows that a selection of foundational research in social psychology must be discarded. At worst this is the tip of the iceberg, and many similar social psychological findings within and without the group polarization tradition will similarly be shown to be unreliable.
---

```{=tex}
\begin{quote}
In our introductory social psychology course, 
we have for many years used the [group polarization experimental paradigm] as
a laboratory exercise. The exercise works beautifully, but one must be
careful to forewarn a class that [group polarization] does not occur with every 
group\ldots and that the effect is not large. 
\par\raggedleft(Brown, 1986, p.\cite[p. 205]{Brown1986})
\end{quote}
```
```{=tex}
\begin{quote}
One of the most robust findings in social psychology is that of attitude polarization 
following discussion with like-minded others.
\par\raggedleft(Cooper et al, 2001, p. 267 \cite[p. 267]{Cooper2001})
\end{quote}
```
## Introduction

Unchecked social and political polarization undermines cooperation and political efficacy [@Mason2018UncivilAgreementBook; @Klein2020]. *Group polarization* is the name for a group-level process that leads to greater opinion extremity in a group as the group deliberates some topic (e.g., "Do you approve, disapprove, or neither approve nor disapprove of the current administration's policies?"). Extremity only increases, though, assuming that the mean opinion of the group is initially biased towards one extreme or the other [@Brown1986; @Brown2020]. Group polarization is important to understand rigorously, therefore, since if groups become increasingly extreme over time, while group opinions are consolidated, this would exacerbate political polarization between political opposition groups, which is already undermines the ability of partisans in democratic governments to work together. Dozens of studies and meta-analyses over decades have apparently established the empirical reality of group polarization [@Moscovici1969; @Myers1970; @Isenberg1986; @Sunstein2009; @Sunstein2019] [@Brown2000, Ch. 5]. We believe group polarization should occur in certain contexts since more extreme group members are likely also to be more stubborn [@Guazzini2015; @Lewandowsky2019], which should pull consensus to extremes [@Turner2020]. However, many or most of the evidence for group polarization relies on experimental designs that use ordinal measurements that are then analyzed as if they were continuous. Unfortunately, this procedure is unfortunately known to induce false positives if it is not properly accounted for [@Liddell2018].

In this paper, we perform a meta-analysis of sixty detections of group polarization from ten published research articles and show that fifty-five of the sixty cannot be reliably identified as group polarization. We show that these 55 are plausibly false detections because they are equally well-explained by a *simple consensus* model, where the mean group opinion is unchanged by group discussion but opinion variance decreases. Theoretically, consensus occurs when sufficiently like-minded individuals influence, which decreases group opinion variance [@DeGroot1974; @Turner2018]. Group polarization, then, is consensus plus an overall increase in group opinion extremity, i.e., an increase in mean group opinion after discussion. We use a novel approach, described herein, that enabled us to identify continuous pre- and post-discussion simple consensus opinion distributions that result in data that appear to be group-polarized when opinions are drawn from these distributions and measured with an ordinal scale. This approach could be adapted for similar social psychological studies that used ordinal measures to determine whether change occurred in distributions of continuous psychological variables such as opinions and beliefs. For social psychology and group polarization, this means that an accepted, heretofore unquestioned, empirical fact must be re-evaluated. Experimental designs in the field must be updated if we are to rigorously understand whether and when group polarization occurs.

In this study, we take the first step towards a new kind of meta-analysis needed to demonstrate that published detections of group polarization are plausibly false. We selected sixty foundational group polarization detections from ten studies. We chose the ten studies guided by a desire to identify foundational studies that contribute to the general consensus in social psychology that group polarization reliably occurs across experimental settings—one author of one of the ten journal articles is a law professor who has written two trade books on group polarization to argue that group polarization occurs in real-world settings in addition to the laboratory, with political and other practical importance [@Sunstein2009; @Sunstein2019]. Our criteria were that the studies were either (1) written by influential authors; (2) had an outsize influence on the study of group polarization in terms of citations and journal prestige; or (3) are well-cited studies that were designed to support one of four recognized theoretical explanations of group polarization. The studies must also have used an ordinal scale for measuring participant opinions (all but a handful of exceptions use ordinal measurements to measure group polarization; e.g. blackjack bets [@Blascovich1973; @Blascovich1976] and mock jury monetary damages [@Schkade2000]). Follow-on studies could establish more objective standards for inclusion, but the above criteria have provided a useful selection of ten diverse studies.

The first prominent group polarization studies of broad scientific interest examined social and political issues without much theoretical development, i.e., lacking specific predictions or explanations. Of these, we include two journal articles, Moscovici and Zavalloni (1969)[@Moscovici1969] who analyzed responses from high school students about domestic and foreign politics and Myers (1970) [@Myers1970] who . The first article is a study of French high school students responding to questions about . The second is \
since group polarization was In tracing the historical development of group polarization, four theoretical explanations have been identified that could possibly be compatible, with some practitioners considering them incompatible. Some influential studies we identify Here we are not interested in evaluating or comparing the these theories, but instead we select influential studies from different theoretical traditions for our selective meta-analysis here.

Each of the ten group polarization studies/journal articles we analyze varied significantly in how groups were formed, how opinions were gathered, and other details. Some

To evaluate whether published detections of group polarization are plausibly false we developed and combined formal and generative models of group polarization to identify plausible simple consensus pre- and post-discussion opinion distributions that seem like group polarization when binned or measured on ordinal scales. Opinion distributions are assumed to be normal. We define a *simple consensus pair* of *latent opinion* distributions to be a pair of pre- and post-discussion distributions where the mean opinion is constant, but the pre-discussion variance is greater than the post-discussion variance. A *latent distribution* of opinions represents the internal, unobservable distribution of opinions for a group or collection of groups. The *observed distribution* is the distribution of reported ordinal opinion measurements. We then define a *plausibly false detection* of group polarization to be one where the mean post-discussion observed distribution is greater in magnitude (i.e., more extreme) than the post-discussion observed distribution's, but both were measurements of a pair of latent simple consensus distributions.

In the formal model, a simulated observed distribution is generated by binning a latent distribution into a finite number of ordinal scale bins, then integrating over the probability density in each bin. In the generative model, an observed distribution is simulated by first drawing participant opinions from a latent distribution, then counting the number of opinions falling within each bin. The formal model is used to identify which, if any, simple consensus pairs exist for a given reported detection of group polarization through an optimization routine that finds pre- and post-discussion variances that generate simulated observations that appear to be group polarization (or reports a failure to such variances). If the optimization routine finds a constant latent mean that generates simulated observed group polarization, then we say the published group polarization detection is plausibly false. If the routine fails to find a constant latent mean that generates simulated group polarization, then we cannot and do not claim the detection is plausibly false. We use the generative model to evaluate how frequently we would expect plausibly false detections when there are finite-size groups, since the formal model implicitly applies infinite-size groups. Therefore in our analyses, we also generate one thousand simulated group polarization datasets for each of the 55 plausibly false detections identified with the formal model and examine the distribution of group polarization effect sizes in terms of Cohen's $d$. In the first round of simulations we set the number of observations to match those in the ten published studies we analyzed. Many or most simulated effect sizes are non-zero with these numbers of observations. To see if perhaps these non-zero effects were due to low statistical power from too few observations, we also tested a hypothetical study with ten times more participants. Finally, we evaluate whether an appropriate statistical model that accounts for the ordinal measurement of latent opinions can reliably identify simple consensus in latent opinions when observed opinions appear as group polarization.

## Methods

## Results

## Discussion

Ordinal scales are an indispensable tool for psychological research on how social influence changes people's beliefs, opinions, knowledge, or similar psychological content. We show here that we must be careful not to mistake the measurement procedure as psychological reality when designing experiments.

Replication crisis, generalizability crisis [@Yarkoni2022], and general lack of coordination in asking the right questions to efficiently understand the mechanisms behind . Understanding mechanisms is the goal of science [@Machamer2000; @Craver2006]. However, in social and psychological sciences, finding mechanisms is more difficult due to the complexity of social psychological systems [@Dubova2023], which require a high density of conceptual definitions that may be difficult to harmonize.