### 2024-04-18 GroupPolarizationStatmod/paper/ms.qmd
More intro content and structure

---
title: "Foundational claims of group polarization are plausibly false, with lessons for the empirical study of psychological dynamics"
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    cite-method: biblatex
    biblatexoptions: sorting=none, style=numeric-comp
    pdf-engine: latexmk
    pdf-engine-opt: -xelatex
    keep-tex: true
    template-partials:
      - _extensions/partials/title.tex
authors:
  - name: Matthew Adam Turner
    affiliation: Stanford University Doerr School of Sustainability, Stanford University
    roles: writing, model development, project development
    corresponding: true
    email: maturner@stanford.edu
  - name: Paul E. Smaldino
    affiliation: 
      - Cognitive and Information Sciences, Univeristy of California, Merced
      - Santa Fe Institute
    roles: writing
    corresponding: true
bibliography: "/Users/mt/workspace/Writing/library.bib"
classoption: abstract
keep-tex: true
geometry:
  - margin=1.1in
  - letterpaper
biblio-title: References
indent: true
suppress-bibliography: true
abstract: |
  \noindent _Group polarization_ is the name for a form of consensus where members of a like-minded group become, on average, more extreme in their opinions after discussing a topic. Group polarization is important because it may drive political polarization if moderate partisans become more extreme. Theoretically, group polarization is celebrated as a regular, reliable outcome of a certain experimental design based on dozens of published studies that purportedly demonstrate the phenomena. However, in this paper we show that 55 of 60 published observations of purported group polarization, from ten foundational studies, are equally well-explained by a simple consensus model where group opinion variance decreases, but extremity does not increase. We use a novel computational meta-analysis framework designed to identify plausible pre- and post-discussion opinion distributions that are consistent with simple consensus, but appear to be group polarization when measured on ordinal, Likert-style scales. This occurs due to ceiling effects induced by ordinal measurements. At best this shows that a selection of foundational research in social psychology must be discarded. At worst this is the tip of the iceberg, and many similar social psychological findings within and without the group polarization tradition will similarly be shown to be unreliable.
---

```{=tex}
\begin{quote}
In our introductory social psychology course, 
we have for many years used the [group polarization experimental paradigm] as
a laboratory exercise. The exercise works beautifully, but one must be
careful to forewarn a class that [group polarization] does not occur with every 
group\ldots and that the effect is not large. 
\par\raggedleft(Brown, 1986, p.\cite[p. 205]{Brown1986})
\end{quote}
```
```{=tex}
\begin{quote}
One of the most robust findings in social psychology is that of attitude polarization 
following discussion with like-minded others.
\par\raggedleft(Cooper et al, 2001, p. 267 \cite[p. 267]{Cooper2001})
\end{quote}
```
## Introduction

Unchecked social and political polarization undermines cooperation and political efficacy [@Mason2018UncivilAgreementBook; @Klein2020]. *Group polarization* is the name for a group-level process that leads to greater opinion extremity in a group as the group deliberates some topic (e.g., "Do you approve, disapprove, or neither approve nor disapprove of the current administration's policies?"). Extremity only increases, though, assuming that the mean opinion of the group is initially biased towards one extreme or the other [@Brown1986; @Brown2020]. Group polarization is important to understand rigorously, therefore, since if groups become increasingly extreme over time, while group opinions are consolidated, this would exacerbate political polarization between political opposition groups, which is already undermines the ability of partisans in democratic governments to work together. Dozens of studies and meta-analyses over decades have apparently established the empirical reality of group polarization [@Moscovici1969; @Myers1970; @Isenberg1986; @Sunstein2009; @Sunstein2019] [@Brown2000, Ch. 5]. We believe group polarization should occur in certain contexts since more extreme group members are likely also to be more stubborn [@Guazzini2015; @Lewandowsky2019], which should pull consensus to extremes [@Turner2020]. However, many or most of the evidence for group polarization relies on experimental designs that use ordinal measurements that are then analyzed as if they were continuous. Unfortunately, this procedure is unfortunately known to induce false positives if it is not properly accounted for [@Liddell2018].

Experimetnal designs in group polarization studies share a certain experimental boilerplate (Figure 1 — NEED TO INCLUDE) with the goal of testing the group polarization hypothesis that a group of like-minded people will develop more extreme opinions on some salient topic over the course of discussing some topic. Group polarization experiments are thus initialized by forming like-minded groups of people and surveying their initial opinions on some topic. In some cases these initial opinions are used for assigning participants to like-minded groups, but other methods have been used, such as geographic location for producing, on average, like-minded groups [@Schkade2010]. Participants then discuss the topic in their groups. After discussion, participant opinions are again measured. Several studies we analyze here determined group polarization occurred (i.e., defined a positive detection) only as a non-zero difference in mean opinion, with no other statistical testing. Some others performed null-hypothesis $t$-tests to test whether there was enough evidence to reject the null-hypothesis that *simple consensus* occurredand not group polarization. In our meta-analysis simulations, we measure simulated group polarization effect size in terms of Cohen's $d$, which is an abstraction beyond these two practices from the group polarization literature that can also be used to compare model fits between metric models that treat group polarization measurements as continuous although they are ordinal, and more appropriate statistical models that account for data ordinality.

In this paper, we perform a meta-analysis of sixty detections of group polarization from ten published research articles and show that fifty-five of the sixty cannot be reliably identified as group polarization. We show that these 55 are plausibly false detections because they are equally well-explained by a *simple consensus* model, where the mean group opinion is unchanged by group discussion but opinion variance decreases. Theoretically, consensus occurs when sufficiently like-minded individuals influence, which decreases group opinion variance [@DeGroot1974; @Turner2018]. Group polarization, then, is consensus plus an overall increase in group opinion extremity, i.e., an increase in mean group opinion after discussion. We use a novel approach, described herein, that enabled us to identify continuous pre- and post-discussion simple consensus opinion distributions that result in data that appear to be group-polarized when opinions are drawn from these distributions and measured with an ordinal scale. This approach could be adapted for similar social psychological studies that used ordinal measures to determine whether change occurred in distributions of continuous psychological variables such as opinions and beliefs. For social psychology and group polarization, this means that an accepted, heretofore unquestioned, empirical fact must be re-evaluated. Experimental designs in the field must be updated if we are to rigorously understand whether and when group polarization occurs.

<!-- In this study, we take the first step towards a new kind of meta-analysis needed to demonstrate that published detections of group polarization are plausibly false. We selected sixty foundational group polarization detections from ten studies. -->

### Studies included in this meta-analysis

We chose the ten studies guided by a desire to identify foundational studies that contribute to the general consensus in social psychology that group polarization reliably occurs across experimental settings—one author of one of the ten journal articles is a law professor who has written two trade books on group polarization to argue that group polarization occurs in real-world settings in addition to the laboratory, with political and other practical importance [@Sunstein2009; @Sunstein2019]. We chose studies that either had a significant influence on the study of group polarization and social psychology, in terms of a high citation count or adaptation in popularized trade books, or were designed to support one of four prominent theoretical explanations of group polarization. The studies must also have used an ordinal scale for measuring participant opinions (all but a handful of exceptions use ordinal measurements to measure group polarization; e.g. blackjack bets [@Blascovich1973; @Blascovich1976] and mock jury monetary damages [@Schkade2000]). We included at least one study that supported or was motivated by one of four established theoretical explanations (introduced below) of group polarization to show all have at least some supporting evidence that must be discarded. We also included two studies that lack a clear theoretical motivation. For the sake of orienting the reader to the group polarization literature and as a useful organizing feature for our analysis, we now introduce the theories that motivate eight of the ten studies. To further orient the reader, we then provide details of the experimental designs used to produce the 60 published positive detections of group polarization across 10 research articles that we scrutinize here. Group polarization experimental designs are diverse and inconsistent with one another in many ways, which complicates this or other potential meta-analyses and may imperil our ability to rigorously understand how regular the group polarization effect really is [@Yarkoni2022; @Almaatouq2022].

The first theoretical explanation whose supporting findings we include here is (1) social comparisons theory, which posits that an individual's opinions are updated so they stand out in a group enough to be "desirably distinctive" (Myers, 1978, p. 562) but not too distinct to be viewed as anti-conformist [@Brown1974; @Sanders1977; @Myers1978]. Second, *persuasive arguments theory* posits that more extreme individuals have a greater number of more convincing arguments for their view, which draw more moderate group members to extreme opinions [@Burnstein1973; @Burnstein1975; @Burnstein1977]. Third, the *self-categorization theory* predicts that group polarization will occur as more moderate individuals judge the mean group opinion to be more extreme [@Turner1989; @Abrams1990; @Krizan2007]. Fourth and finally, the social decision schemes explanation posits that the structure of [@Davis1973; @Zuber1992; @Friedkin1999a]. We do not evaluate or compare these theories here, though a simpler explanation that extremists are more stubborn may be a more parsimonious explanation consistent with these four alternatives [@Acemoglu2013; @Guazzini2015; @Turner2020].

Each of the ten group polarization studies we analyze here varied significantly in how groups were formed, how opinions were gathered, and other details. There was significant variation across experimental designs, participant population sizes, group sizes, and how like-minded groups were formed in the first place. We review some of these details below to provide the reader with some insight into the experimental designs that produced what we show are plausibly false positive detections of group polarization. We focus here on the questions that were asked, how like-minded novel experimental groups were formed, organizing our presentation by theoretical tradition.

#### Details of experimental conditions included in this meta-analysis

\paragraph{Atheoretical} The first of ten studies we included in our meta-analysis was by Moscovici and Zavalloni (1969) [@Moscovici1969] who sought to measure group polarization among groups of French high school students who discussed domestic and foreign politics. Moscovici and Zavalloni tested four conditions whether group polarization occurred for two different questions and two different rating systems: participants were asked their opinions on then-president Charles De Gaulle and their opinion about foreign aid from the USA either in the "Opinion" setting or the "Judgement" condition (see Table 4 on p. 132).

The second, relatively more recent article is Schkade, Sunstein, and Hastie (2010) [@Schkade2010] who used geographic location to construct like-minded groups by running a group polarization experiment in two cities in the state of Colorado in the United States: Colorado Springs, whose residents historically voted for republican candidates, and Boulder, whose residents historically voted for democratic candidates. Discussion groups were formed in which they measured group polarization when group members discussed global warming, affirmative action, and same-sex civil unions.

\paragraph{Social comparisons} In developing the social comparisons theory of group polarization (described in Myers (1978) [@Myers1978]), Myers and Bishop (1970), writing in *Science,* [@Myers1970] sought to measure group polarization among high school students in the United States who discussed racial attitudes and Myers (1975) [@Myers1975] measured group polarization in groups of college students discussing the quality of a faculty member in one condition, and women's rights and issues in another condition.

\paragraph{Persuasive arguments}

\paragraph{Self-categorization}

\paragraph{Social decisions scheme}

The ten studies we included each tested several experimental conditions meant to elicit group polarization in novel experimental groups, 60 in total, which we briefly introduce here. \

### Research overview

To evaluate whether published detections of group polarization are plausibly false we developed and combined formal and generative models of group polarization to identify plausible simple consensus pre- and post-discussion opinion distributions that seem like group polarization when binned or measured on ordinal scales. Opinion distributions are assumed to be normal. We define a *simple consensus pair* of *latent opinion* distributions to be a pair of pre- and post-discussion distributions where the mean opinion is constant, but the pre-discussion variance is greater than the post-discussion variance. A *latent distribution* of opinions represents the internal, unobservable distribution of opinions for a group or collection of groups. The *observed distribution* is the distribution of reported ordinal opinion measurements. We then define a *plausibly false detection* of group polarization to be one where the mean post-discussion observed distribution is greater in magnitude (i.e., more extreme) than the post-discussion observed distribution's, but both were measurements of a pair of latent simple consensus distributions.

In the formal model, a simulated observed distribution is generated by binning a latent distribution into a finite number of ordinal scale bins, then integrating over the probability density in each bin. In the generative model, an observed distribution is simulated by first drawing participant opinions from a latent distribution, then counting the number of opinions falling within each bin. The formal model is used to identify which, if any, simple consensus pairs exist for a given reported detection of group polarization through an optimization routine that finds pre- and post-discussion variances that generate simulated observations that appear to be group polarization (or reports a failure to such variances). If the optimization routine finds a constant latent mean that generates simulated observed group polarization, then we say the published group polarization detection is plausibly false. If the routine fails to find a constant latent mean that generates simulated group polarization, then we cannot and do not claim the detection is plausibly false. We use the generative model to evaluate how frequently we would expect plausibly false detections when there are finite-size groups, since the formal model implicitly applies infinite-size groups. Therefore in our analyses, we also generate one thousand simulated group polarization datasets for each of the 55 plausibly false detections identified with the formal model and examine the distribution of group polarization effect sizes in terms of Cohen's $d$. In the first round of simulations we set the number of observations to match those in the ten published studies we analyzed. Many or most simulated effect sizes are non-zero with these numbers of observations. To see if perhaps these non-zero effects were due to low statistical power from too few observations, we also tested a hypothetical study with ten times more participants. Finally, we evaluate whether an appropriate statistical model that accounts for the ordinal measurement of latent opinions can reliably identify simulated simple consensus in latent opinions when simulated observations appear as group polarization based on mean group opinion alone.

## Methods

## Results

## Discussion

Ordinal scales are an indispensable tool for psychological research on how social influence changes people's beliefs, opinions, knowledge, or similar psychological content. We show here that we must be careful not to mistake the measurement procedure as psychological reality when designing experiments.

Replication crisis, generalizability crisis [@Yarkoni2022], and general lack of coordination in asking the right questions to efficiently understand the mechanisms behind . Understanding mechanisms is the goal of science [@Machamer2000; @Craver2006]. However, in social and psychological sciences, finding mechanisms is more difficult due to the complexity of social psychological systems [@Dubova2023], which require a high density of conceptual definitions that may be difficult to harmonize.

# Tables {.unnumbered}

asdf