### 2025-06-19 GroupPolarizationStatmod/paper/intro.qmd
init intro.qmd for only intro drafting

---
title: "Group Polarization Revisited"
author: "Your Name"
format:
  pdf:
    documentclass: article
    keep-tex: true
bibliography: 
  - this.bib
  - /Users/mt/workspace/Writing/library.bib
output: pdf
---

# Introduction

Group polarization is one of the most widely cited but unevenly evidenced phenomena in social psychology...

*Write like you mean it.*  
*You're not drafting — you're building.*