### 2025-08-17 GroupPolarizationStatmod/paper/intro-latest.tex
expanded discussion and intro

\subsection*{Abstract}\label{abstract}
\addcontentsline{toc}{subsection}{Abstract}

\emph{Group polarization} is the name for a form of consensus where
members of a like-minded group become, on average, more extreme in their
opinions after discussing a topic. Group polarization is important
because it may increase social and political tensions if moderates
become more extreme. Decades of studies have replicated detections of
group polarization, typically using an ordinal scale to measure
opinions. However, these studies did not account for ordinal
measurements, which can result in spurious detections of opinion change.
Lacking original data we can still calculate the probability that a
published detection of group polarization was spurious, i.e., to
calculate the \emph{false detection rate}. Across 54 group polarization
experimental conditions from ten representative journal articles, we
calculated false detection rates to be between 0.52 and 0.88, with a
median of 0.75, using a generative model of group polarization
experiments seeded with empirical data. We also use our model to develop
experimental designs that achieve an acceptable false discovery rate.
Much of group polarization research may be unreliable. This analysis can
help change that by enabling others to avoid this hidden pitfall. More
broadly this work demonstrates one important way that replication
success alone does not imply epistemic reliability.

\section{Introduction}\label{introduction}

\begin{quote}
In our introductory social psychology course, we have for many years
used the {[}group polarization experimental paradigm{]} as a laboratory
exercise. The exercise works beautifully, but one must be careful to
forewarn a class that {[}group polarization{]} does not occur with every
group\ldots and that the effect is not large (Brown 1986).
\end{quote}

\begin{quote}
One of the most robust findings in social psychology is that of attitude
polarization following discussion with like-minded others (Cooper,
Kelly, and Weaver 2001).
\end{quote}

\begin{quote}
  The good thing about science is that it's true, whether or not you believe
  it (Neil de Grasse Tyson t-shirt, 2010's).
\end{quote}

If an extremist's opinion falls and it is measured with a Likert scale, will
it make a noise? The answer had better be yes if we want to properly measure
opinion change. Accurate measurements of opinion change when tracking
opinions and beliefs over time in surveys and experiments is critical to
design more, sustainable, responsive, and efficient governments,
institutions, and policies (Mason 2018; Klein 2020). One important social
psychology paradigm for explaining opinion change is called \emph{group
polarization}.  Former Obama White House administrator of Information and
Regulatory affairs, and current Harvard Law School professor, Cass Sunstein
writes that the scientific ``law of group polarization''~\cite{Sunstein1999}
provides a ``clue'' to explaining how ``facism\ldots student
radicalism\ldots Islamic terrorism\ldots{[}t{]}he Rwandan
genocide\ldots{[}e{]}thnic Conflict\ldots acts of torture and humiliation by
American soldiers at Abu Ghraib'', and more \cite[p. 1]{Sunstein2009}.

We report here that, unfortunately, many studies of group polarization 
used an ordinal, Likert-style, measurement scale but failed to account 
for this fact.  
Naive measurement procedures like this are known to induce
both false positive and negative detections of changes in opinion, 
belief, or similar \emph{latent} psychological entities, i.e., unobservable
theoretical variables~\cite{Liddell2018,Turner2020}.
Ordinal scales are ordered, integer-valued responses that represent both
degree and polarity of opinion, e.g., ~-3
indicates ``Strongly disagree'', +3 indicates ``Strongly agree'', 0
is neutral, with intermediate numbers indicating different degrees of
disagreement or agreement, depending on the sign.

To answer the question posed above in the paper's lead sentence, it is
entirely possible that ordinal scales will not detect all changes in extreme
opinions. For example, consider the 7-point Likert scale described above. 
If an extremist's latent opinion changes
from -5 to -4, but they map their latent opinion onto the closest integer
option, they will report -3 each time: such opinion shifts towards
moderation would not be detected. On the other hand, a latent opinion shift
from -1 to -2 would be detected. 
Considering a group composed of two individuals whose opinions changed like 
that, their mean latent opinion did not change: it is -3 in both the pre-
and post-deliberation phases. We call this \emph{simple consensus}, where
individuals' opinions agreed more following deliberation, but the mean
opinion stayed constant. In statistical terms, the \emph{mean} opinion
stayed the same and the \emph{variance} decreased. 
The 7-point Likert scale, however, spuriously measures a change in 
mean opinion from $\frac{1}{2}(-3 + -1) = -2$ to $\frac{1}{2}(-3 + -2) =
-2.5$. In other words, we have a false detection of group polarization where
the mean opinion seems to shift from -2 to -2.5.  The same logic applies to
groups with more than two people.

We developed a generative model of opinion change to identify, if possible,
two opinion distributions representing \emph{simple consensus} that seem to
be \emph{group polarization} when transformed to an ordinal measurement
scale.  latent distributions that could spuriously appear like group
polarization.  This type of problem is known to happen when ordinal
measurements are taken of continuous data without statistically accounting
for the measurement procedure itself~\cite{Liddell2018}.  We further applied
our method to estimate a best-case probability that a published detection of
group polarization is false, across 57 experimental conditions in a corpus
of ten journal articles curated for their prominence and representativeness
of group polarization research.  We turn this criticism of past blunders
constructive by providing a procedure to calculate what Cohen's $d$ should
be used to guarantee a family-wise error rate (i.e., Type I error rate) of
5\% or less.

We calculate the severity of the problem in two steps. In the first step we
provide proof by example that there exists a null opinion change model in
terms of \emph{latent, continuous} psychological opinions that looks like
group polarization when measured via participant behavior, i.e., by their
response on an \emph{ordinal survey scale}. If such a null model is found,
this means that we cannot rule out a null effect, given the published data.
This means, therefore, that the published detection is just as likely false
as it is true, so we say it is a \emph{plausibly false detection of group
polarization}, or a \emph{plausibly spurious detection} for short.

In the second step of the analysis of whether, we simulate observations of
group polarization experiments where participant opinions are drawn from the
null model distributions identified in the first step, then these are
measured on a Likert scale for simulated group polarization survey responses
on an ordinal scale. We then calculate the family-wise error rate (i.e.,
Type I error rate) and false detection rate using various effect sizes to
determine significance of the effect in terms of Cohen's $d$.  The
family-wise error rate is defined as the probability that a known null
effect is detected as significant. We use this to then estimate the *false
detection rate*, which is the probability that a detection of an effect is a
false one, which depends on the relative prevalence of true and false
positive detections in the literature.  Finally, we put our model to
constructive use to show what Cohen's $d$ must be set to to limit the
family-wise error rate to 5\%, denoted $d*$. This will not help rescue any
of the published results here found to be plausibly spurious. But, it
provides a measure of the weakness of the published experimental designs
independent of their shared fatal failure to account for the measurement
procedure: greater $d*$ indicates weaker designs.

We find that 95\% (54 of 57) of group polarization detections are plausibly
spurious.  In other words, empirical support for group polarization has been
seriously undermined just by removing these results from the count of group
polarization replications.  There is no reason to expect other papers to be
more reliable. (Summarize other two findings, 1. FWER/FDR and 2. $d*$ for
FWER < 5\%)

Group polarization is not the only phenomenon in the social and behavioral
sciences to naively ignore measurement procedures.  Liddell and Kruschke
found that the measurment procedures of experimental designs are rarely, if
ever, accounted for by social psychology researchers in their systematic
review of select social psychology journals. Our work is important, then,
both to develop rigorous group polarization studies and provide a template
for re-analyzing pre-crisis social and behavioral research.  Our methods can
be turned on any data-driven social science analysis of data where
participants report a latent, subjective psychological variable (mood,
physiological state, etc.), and where various theories supposedly battle it
out to best represent the phenomenon of interest via null-hypothesis
significance testing using ordinal data as continuous.  This may just be the
first of many bricks to fall since \emph{all 68} of the articles that
mentioned the word ``Likert'' and analyzed ordinal data used metric models
to detect latent psychological variable change across three influential
psychology journals, \emph{Psychological Science}, the \emph{Journal of
Experimental Psychology: General}, and the \emph{Journal of Personality and
Social Psychology} (Liddell and Kruschke 2018). This further confirms that
replication is far from the only foundational crisis that behavioral science
faces\cite{Yarkoni2022,Turner2022}.

\subsection{Group polarization research and corpus}

To situate this current paper it will help to introduce some relevant
research traditions and common experimental designs in group polarization,
using the ten studies included in our corpus as touchstones. To begin,
*group polarization* is defined as...(continue basic definition with some
examples from our corpus).

Paragraph dos: Common experimental, measurement, and statistical procedures
to induce and detect group polarization

\subsection{Failure to account for the measurement process can induce
false detections} \label{ssubsec:false-detections}

To see how failing to account for the measurement procedure could be a
problem, let's step back to sketch a \emph{model} of group polarization
experimental design.  This just means we are going to pick some key
components and processes that we assume are important together. This does
not mean other factors are not important.  We do this to focus on one
complex system by \emph{screening out} interactions with several other
complex systems. This piecemeal process may seem slow, but it has proven
useful in more mature sciences like physics and biology.

Group polarization experiments share a common experimental design where,
first, participant opinions are gathered on a topic of interest. Next,
groups are formed that have some initial opinion extremism. Participants
deliberate or discuss the topic for some set time. After the set time,
participants again give their opinion. Group polarization is said to occur
if the mean group opinion became more extreme, i.e., if the magnitude of the
mean group opinion is greater \emph{post-deliberation} compared to
\emph{pre-deliberation}. For example, \cite{Moscovici1969} asked Parisian
high school students to indicate their degree of agreement with the
statement, ``American economic aid is always used for political pressure''.
There were seven options in the selection set, -3 indicating ``strongly
disagree'', +3 indicating ``strongly agree'', and 0 was neutral. Moscovici
and Zavalloni report that the mean pre-deliberation opinion was -0.61 and
the mean post-deliberation opinion was -1.09 (Moscovici and Zavalloni, 1969;
Table 4, p. 132). This is a -0.48 shift to more extreme disagreement with
the statement, reported by Moscovici and Zavalloni as statistically
significant in Table 5 on p. 132. Sometimes group polarization research uses
significance testing for determining group polarization, but Moscovici and
Zavalloni, and others, just report a bare difference in mean opinion. This
alone should make modern researchers think twice before including it and
others as canon, but apparently this has gone unnoticed.

Note that only \emph{behaviors} are measured in this experiment, and not
\emph{opinions}.  Instead, the participant (paid by the job, not the hour)
navigates their mouse to hover over a radio button in an online survey, or
voices their opinion by phone to a survey caller. The \emph{latent}, complex
neural activity that underlies opinion formation and expression via
different behaviors is unseen and, so far, unknown. We call internal
opinions \textit{latent opinions}, from the Latin \emph{latentem} meaning
``lie hidden''.  In other words, we cannot measure neurobiological activity
and infer a participant's opinion.  We only ever measure participant
behavior that signals the participant's opinion. Note this presupposes that
the participant's opinion is that behavior best matches their opinion on the
deliberation topic.

The problem arises when we try to compare average \emph{measured} opinion
between phases, when instead the \emph{latent} opinion is the one that would
prove group polarization occurred.  In fact, the problem seems to be that
they are assumed to be identical. However, group polarization of latent
opinions, or other psycho-social dynamics, are known to get distorted when
measured with an ordinal scale~\cite{Liddell2018,Turner2020}. Recall that
theoretically group polarization is one type of consensus, where extremism
and agreement both increase, meaning the magnitude of the mean opinion in
the group increases while the variance decreases. We distinguish group
polarization from what we call \emph{simple consensus}

of If we represent the latent distribution

(PREVIOUS TWO PARAGRAPHS SET UP THIS FINAL PARAGRAPH: CONTINUE TO USE
MOSCOVICI 1969 TO EXPLAIN IN THE SIMPLEST TERMS HOW FAILING TO ACCOUNT FOR
MEASUREMENT PROCEDURE CAN LEAD TO FALSE POSITIVES; see Figure
\ref{fig:distros})

\subsection{Overview}

In the next section, we introduce the various formal and computational
Methods used to obtain our findings outlined above. We then explain our
findings in detail in the Analysis section. We close with a Discussion of
how this is not a defeat, but a new beginning for group polariation, given
that there are good reasons to expect group polarization to occur. We also
explain how to expanding this analysis to examine other published studies on
group polarization and across social and behavioral science disciplines.