### 2025-10-30 GroupPolarizationStatmod/paper/article-table.tex
added article table

\begin{table}[ht]
\centering
\caption{\textbf{Journal articles in our corpus.}}
\label{tab:articles}

\begin{tabular}{m{0.9in} m{1.35in} m{2.45in} m{1.3in}}
\toprule
\textbf{Category} & \textbf{Authors (Year)} & \textbf{Title} & \textbf{Journal} \\
\midrule

\rowcolor{gray!10}
\multirowcell{2}{Politics} 
                  & Moscovici \& Zavalloni (1969)~\cite{Moscovici1969} & The Group as a Polarizer of Attitudes & \emph{JPSP} \\
\rowcolor{gray!10}
                  & Schkade, Sunstein, \& Hastie (2010)~\cite{Schkade2010} & When Deliberation Produces Extremism & \emph{Critical~Review} \\[0.2em]

\rowcolor{gray!5}
\multirowcell{2}{Social Comp's \vspace{0.7em}}
% \multirowcell{2}{Social Comp's}
                  & Myers \& Bishop (1970)~\cite{Myers1970} & Discussion Effects on Racial Attitudes & \emph{Science} \\[0.2em]
\rowcolor{gray!5}
                  & Myers (1975)~\cite{Myers1975} & Discussion-Induced Attitude Polarization & \emph{Human Relations} \\[0.3em]

\rowcolor{gray!10}
\multirowcell{2}{Persuasive Arg's}
                  & Burnstein \& Vinokur (1973)~\cite{Burnstein1973} & Testing Two Classes of Theories About Group Induced Shifts In Individual Choice & \emph{JESP} \\
\rowcolor{gray!10}
                  & Burnstein \& Vinokur (1975)~\cite{Burnstein1975} & What a person thinks upon learning he has chosen differently from others: Nice evidence for the persuasive-arguments\ldots & \emph{JESP} \\

\rowcolor{gray!5}
\multirowcell{3}{Self-Categ'n}
                  & Abrams, et al., (1990)~\cite{Abrams1990} & Knowing What to Think by Knowing Who You Are: Self‐Categorization and the Nature of Norm Formation\ldots & \emph{Brit. J. of Soc. Psy.} \\
\rowcolor{gray!5}
                  & Hogg, Turner, \& Davidson (1990)~\cite{Hogg1990} & Polarized Norms and Social Frames of Reference: A Test of the Self-Categorization Theory of Group Polarization & \emph{B. \& Appl. Soc. Psy.} \\
\rowcolor{gray!5}
                  & Krizan \& Baron (2007)~\cite{Krizan2007} & Group Polarization and Choice Dilemmas: How Important is Self-Categorization? & \emph{Eur. J. of Soc. Psy.} \\

\rowcolor{gray!10}
\multirowcell{1}{Soc'l Decisions}
                  & Friedkin (1999)~\cite{Friedkin1999} & Choice Shifts \& Group Polarization & \emph{Amer. Soc. Rev.} \\

\bottomrule
\end{tabular}
\end{table}