### 2025-11-14 GroupPolarizationStatmod/paper/method.tex

\section{Method}\label{method}

A group polarization replication is proven undecidable, and therefore not
truly a replication, if we can find a model of \emph{mere agreement} that
appears to be polarized under that replication's experimental design. If we
cannot find such a model, we cannot determine a replication to be
undecidable. Mere agreement in this model is a dynamic where agreement
increases within the group, but extremism does not—in other words, A
\emph{null-polarization distributions} is a set of two Gaussian
distributions: one representing the pre-discussion opinions of the group, and
the other representing the post-discussion opinions. Our
\emph{null-polarization distributions} represent the process of the group
finding \emph{mere agreement}, which is when the average group opinion is
constant over time, but opinion variance decreases to represent agreement
(or, equivalently, \emph{consensus}~\cite{DeGroot1974}).

Null-polarization  We thus model opinions at each time, $t=\mathrm{pre}$ and
$t=\mathrm{post}$, as random variables drawn from a normal distribution with
mean $\mu_t$ and standard deviation $\sigma_t$.  we drop the time

The null-polarization distributions are represented by

in a simple, general way—i.e., a decrease in opinion variance but a static
opinion mean—we represent . In the general group polarization hypothesis, the
group is assumed to start out mostly agreeing in principle (i.e., mostly all
agree or disagree) but to different extremes. After discussion

Showing a replication is undecidable is sufficient to undermine it. However,
it does not tell us what is the probability that a particular experimental
design produces false positives, due to clipping or any other problem.

first develop a mathematical model of estimate the best-case scenario false
positive rate for each experimental design,

\subsection{Mathematical Model of Null-Polarization Measurement Artifacts}

Latent opinions are modeled as normally-distributed random variables.  
This is a simple, sufficiently general way to represent group-level changes
in opinions as encoded in the general hypothesis of group polarization, that
likeminded groups tend to become more extreme on average as they discuss.
For time, $t$, then, we represent each participant $i$'s opinion as 
\begin{equation} 
  \omega_{i,t} \sim \mathcal{N}(\mu_t, \sigma_t).
  \label{eq:opinionDistribution} 
\end{equation} 
\noindent

Group-level opinion change, then, can be represented as a change in opinion
mean, opinion variance, or both. Group polarization occurs 
when a group's average opinion increases over time. When likeminded people
interact, they tend to come to agree more over time—they \emph{find consensus}. 
By definition this means opinion variance decreases, 
whether the average changes or not.

We can define a null-polarization model, then, as one where the group's
average opinion stays constant, but the variance decreases: $\sigmapost
\leq \sigmapre$. Since the mean opinion is constant we just call it 
$\mu = \mupre = \mupost$. For conciseness, we define a null-polarization
model as a 3-tuple, $\nu = (\mu, \sigmapre, \sigmapost)$.

\input{method-tables.tex}

\subsubsection{Simulated opinion measurements}

We simulate the collection of an \emph{ordinal distribution} of \emph{ordinal
observations}, as was performed in a prototypical group polarization experiment 
by integrating the latent opinion probability density function, $f(\omega)$,
over each bin's corresponding range of opinion values. For example, consider an
ordinal scale that runs from -3 (strongly disagree) to +3 (strongly agree)
(as used by \textcite{Moscovici1969}, for example).
the frequency with which the ``2'' bin is chosen is just the total
area underneath the probability density curve between 1.5 and 2.5.

Formally, the probability density function is
\begin{equation}
  f(\omega) = 
    \frac{1}{\sigma \sqrt{2 \pi}} e^{-\frac{(\omega - \mu)^2}{2 \sigma^2}}.
  \label{eq:normal_pdf}
\end{equation}
\noindent
Assume there are $B$ bins, meaning there are $B+1$ \emph{thresholds} that
define which latent opinions get mapped to which bin values. The bin value,
$b_i$, is the numerical value of the $i$th bin.

\subsection{Root-finding Algorithm to Find Null-Polarization Models with Artifacts}

\subsection{Corpus Analysis to Determine Undecidable Replications}

\subsection{Evaluating Grouop Polarization Experimental Designs}