### 2025-11-19 GroupPolarizationStatmod/paper/code_examples.qmd
Created paper/code_examples.qmd and continued on method

---
title: Code examples to accompany "If the Null Fits, You Must Omit"
author: Matt Turner · [matt@mat.phd](mailto:matt@mat.phd)
editor: 
  markdown: 
    wrap: 90
bibliography: this.bib
---

# Purpose: demonstrate gp-stats codebase

Here are some quick notes on the group polarization codebase in R supporting our article,
*If the Null Fits, You Must Omit*. The code is permanently [hosted on
GitHub](https://github.com/mt-digital/gp-statmod/tree/final-draft/GroupPolarizationStatmod).

To follow along, please clone the `gp-statmod` directory.

```         
$ git clone https://github.com/mt-digital/gp-statmod.git
```

Change directories to go to the `paper` directory where this document lives:

```         
cd gp-statmod/GroupPolarizationStatmod/paper
```

Then you can run the examples in this document, edit them, etc., on your own computer.

## Calculate probability density vector for `Moscovici1969_Americans` condition

See Method section of the paper where we calculate the probabilty density vector for
different measurements based on the latent opinion probability density function.

Here we show how to calculate the mean observed opinion at pre- and post-discussion to
show how spurious group polarization occurs under a null polarization hypothesis.

@Moscovici1969 report observing a mean opinion of -0.6 in pre-discussion opinions and mean
opinion -1.04 post-discussion, measured using an ordinal scale from -3 (strongly disagree)
to +3 (strongly agree). Let's calculate the probability density vector for the pre- and
post-discussion distributions then calculate the mean opinion for each one. The latent
distributions have the same mean, -1.2, but different variances, which results in a
shifted observed mean due to opinion clipping.

![](images/clipboard-1443885641.png){width="628"}

We call this a *null polarization model* (defined in terms of latent opinions) that
generates *spurious group polarization.* *Spurious group polarization* is when the
observed opinion distributions appear to polarize (i.e., *radicalize*, or *become more
extreme*) that does not represent the true change in latent opinions. The measurement
procedure *clips* extreme opinions less than the minimum bin value, or greater than the
maximum, causing the average observed opinion to *spuriously* appear polarized.

Let's use some modeling code from the repository to demonstrate. For this part we need to
source (*use code from*) the `model.R` file that has the the
[`binProb`](https://github.com/mt-digital/gp-statmod/blob/final-draft/GroupPolarizationStatmod/model.R#L7)
function I use below. Also make sure you have set the working directory to the
`GroupPolarizationStatmod` directory within the `gp-stats` project/repository directory.

For a preliminary step, adapt my code below for your computer. Change the argument to
`root.dir` to be the location of the `GroupPolarizationStatmod` directory on your own
computer.

```{r setup}

```

Now in the following chunk, the `source` function will know where to look for the file
with the `makeProbVec` function we need, `model.R`.

```{r warning=FALSE}
# This is how I set the working directory
# Adapt for your own computer
setwd("~/workspace/gp-statmod/GroupPolarizationStatmod/")
# Load makeProbVec from file model.R
source("model.R")
```

Now read in the pre- and post-discussion means from `Moscovici1969_Americans`

```{r}
# Experimental means
mosc1969_ave_pre <- -0.6
mosc1969_ave_post <- -1.04
```

Create a hypothesized null model that will spuriously look like polarization, with the
same pre- and post-discussion means above. These correspond to $\mu$,
$\sigma_\mathrm{pre}$, and $\sigma_\mathrm{post}$ from the paper that define a null
polarization model, $\nu = (\mu, \sigma_\mathrm{pre}, \sigma_\mathrm{post})$.

```{r}
# Hypothesized null model that looks polarized
mu <- -1.2
sd_pre <- 4.3
sd_post <- 1.9
```

These

```{r}
# Vector w/ probability density for each bin
f_pre <- makeProbVec(-3:3, -1.2, 4.3)
# Print to two significant digits
sprintf("%.2f", f_pre)
```

We can plot this simply using just

```{r}
barplot(f_pre, names.arg = -3:3)
```