### 2025-11-20 GroupPolarizationStatmod/paper/method.qmd

---
editor: 
  markdown: 
    wrap: 100
---

# Method

A group polarization replication is proven undecidable, and therefore not truly a replication, if we
can find a model of *mere agreement* that appears to be polarized under that replication's
experimental design. If we cannot find such a model, we cannot determine a replication to be
undecidable. Mere agreement in this model is a dynamic where agreement increases within the group,
but extremism does not---in other words, A *null-polarization distributions* is a set of two
Gaussian distributions: one representing the pre-discussion opinions of the group, and the other
representing the post-discussion opinions. Our *null-polarization distributions* represent the
process of the group finding *mere agreement*, which is when the average group opinion is constant
over time, but opinion variance decreases to represent agreement (or, equivalently,
*consensus* [@DeGroot1974]).

Null-polarization We thus model opinions at each time, $t=\mathrm{pre}$ and $t=\mathrm{post}$, as
random variables drawn from a normal distribution with mean $\mu_t$ and standard deviation
$\sigma_t$. we drop the time

The null-polarization distributions are represented by $\nu$.

in a simple, general way---i.e., a decrease in opinion variance but a static opinion mean---we
represent . In the general group polarization hypothesis, the group is assumed to start out mostly
agreeing in principle (i.e., mostly all agree or disagree) but to different extremes. After
discussion

Showing a replication is undecidable is sufficient to undermine it. However, it does not tell us
what is the probability that a particular experimental design produces false positives, due to
clipping or any other problem.

## Simulation of Mere Consensus and Spurious Polarization

Our analysis models \emph{mere consensus} to examine whether published detections of group polarization may plausibly be \emph{spurious}, caused instead by artificats introduced by opinion measurement design.

Latent opinions are modeled as normally-distributed random variables. This is a simple, sufficiently
general way to represent group-level changes in opinions as encoded in the general hypothesis of
group polarization, that likeminded groups tend to become more extreme on average as they discuss.
For time, $t$, then, we represent each participant $i$'s opinion as 
$$
\omega_{i,t} \sim \mathcal{N}(\mu_t, \sigma_t).
$$

Group-level opinion change, then, can be represented as a change in opinion mean, opinion variance,
or both. Group polarization occurs when a group's average opinion increases over time. When
likeminded people interact, they tend to come to agree more over time---they *find consensus*. By
definition this means opinion variance decreases, whether the average changes or not.

We can define a null-polarization model, then, as one where the group's average opinion stays
constant, but the variance decreases: $\sigmapost
\leq \sigmapre$. Since the mean opinion is constant we just call it $\mu = \mupre = \mupost$. For
conciseness, we define a null-polarization model as a 3-tuple, $\nu = (\mu, \sigmapre, \sigmapost)$.

::: {style="text-align: center;"}
\[ TABLE \[tab:latent-variables\](#tab:latent-variables){reference-type="ref"
reference="tab:latent-variables"} ABOUT HERE \]
:::

### Simulated group polarization measurements

We simulate the collection of an *ordinal distribution* of *ordinal observations*, as was performed
in a prototypical group polarization experiment by integrating the latent opinion probability
density function, $f(\omega)$, over each bin's corresponding range of opinion values. For example,
consider an ordinal scale that runs from -3 (strongly disagree) to +3 (strongly agree) (as used by
@Moscovici1969, for example). the frequency with which the "2" bin is chosen is just the total area
underneath the probability density curve between 1.5 and 2.5.

Formally, the probability density function is $$
f(\omega) = 
    \frac{1}{\sigma \sqrt{2 \pi}} e^{-\frac{(\omega - \mu)^2}{2 \sigma^2}}.
$$ To set up our measurement simulation, we will represent oridnal bin values as $b_i$ with
$i=1,\ldots,B$---i.e., there are $B$ bins. To create $B$ bins there must be $B+1$ *thresholds* that
define what range of latent opinions correspond to which ordinal bin value.

::: {text-align="center"}
\[ TABLE [\[tab:experiment-parameters\]](#tab:experiment-parameters){reference-type="ref"
reference="tab:experiment-parameters"} ABOUT HERE \]
:::

All bins have a width of 1 in latent opinion space, except for the two most extreme bins at opposite
ends of the sentiment spectrum---this works because bin values are always separated by 1. All
continuous latent opinion values $\omega < b_1 + \frac{1}{2} = \theta_1$ get mapped to the ordinal
measurement value $b_1$---so, $b_1$'s lower threshold is $\theta_0 = -\infty$. Similarly, all
opinions $\omega > b_B - 0.5 = \theta_B$ are reported as the maximum bin value, $b_B$, with an
effective upper bound of $\theta_{B+1} = +\infty$. We can define all thresholds compactly as
[@Liddell2018] $$
\theta_j = \begin{cases}
  -\infty         & \text{ if } j=0 \\
  \infty          & \text{ if } j=B \\
  b_j + \frac{1}{2} & \text{ otherwise.}
\end{cases}
$$

We model the measurement of ordinal opinions by calculating the probability density function over
bins, written $$
f[b_i] = \int_{\theta_{i - 1}}^{\theta_i} f(\omega) d\omega,
$$ switching to square brackets for the argument to $f$ now to emphasize that bins are discrete.
With this ordinal measurement model we can caluclate the mean opinion of the simulated measurements
like so $$\bar{o}_t = \frac{1}{B} \sum_{b=b_1}^{b_B} b \cdot f_t[b].$$

Group polarization is the difference between the post-deliberation mean observed opinion and the
pre-deliberation one, which we can write as $$g = \bar{o}_\mathrm{post} - \bar{o}_\mathrm{pre}.
  \label{eq:group-polarization}$$

::: center
\[ TABLE [\[tab:simulated-observations\]](#tab:simulated-observations){reference-type="ref"
reference="tab:simulated-observations"} ABOUT HERE \]
:::

### Simulated spurious group polarization

Group polarization measures radicalization in experimental echo chambers in terms of the difference
between mean group opinions before and after the group discusses some survey topic. The following
model shows the problem of how the measurement scheme above can make *mere agreement* spuriously
look like polarization.

To make things more exact, we define a *null model*, $\nu$, of mere agreement defined as a 3-tuple:
a constant mean for the pre- and post-discussion latent opinion distributions, and a
pre-deliberation latent opinion variance that is greater than the post-deliberation variance to
represent finding agreement, $$
\nu = (\mu,~\sigmapre,~\sigmapost).
$$ Group polarization is *spurious* in this case if measurement clipping causes the observed
difference in pre- and post-discussion opinions to be greater than zero, i.e., $g > 0$ (Equation
[\[eq:group-polarization\]](#eq:group-polarization){reference-type="ref"
reference="eq:group-polarization"}, even if the true latent change is zero.

We will show how this arrangement generates spurious polarization in an experimental condition in
our corpus, the *Americans* experimental condition in @Moscovici1969, to which we gave the
Experiment ID `Moscovici1969_Americans`. In that condition, the authors used a seven-bin scale
($B=7$) with numerical values from -3 to +3 indicating opinions from "strongly disagree" to
"strongly agree". Pre-consensus opinions were reported to have mean -0.6 and the post-consensus mean
was -1.04 [@Moscovici1969 p. 132, Table 4].

And now we will show that a null model with
$\nu = (\mu = -1.2, \sigmapre = 4.3, and \sigmapost = 1.9)$ generates spurious polarization with
exactly the pre- and post-discussion means Moscovici and Zavalloni report
(Figure [\[fig:distros\]](#fig:distros){reference-type="ref" reference="fig:distros"}). First,
calculate the vector of probability density values for each bin, $f[b]$, for the pre- and
post-discussion latent opinion distributions using Equation [\[eq:\]](#eq:){reference-type="ref"
reference="eq:"} (calculated with
\['makeProbVec'\](https://github.com/mt-digital/gp-statmod/blob/final-draft/GroupPolarizationStatmod/model.R#L43)
from the \[project code
repository\](https://github.com/mt-digital/gp-statmod/tree/final-draft/GroupPolarizationStatmod)

$$
    f_\mathrm{pre}[b] = 
$$ $$
    f_\mathrm{post}[b] = 
$$

::: {style="text-align: center;"}
FIGURE [\[fig:distros\]](#fig:distros){reference-type="ref" reference="fig:distros"} ABOUT HERE
:::

## Root-finding Algorithm to Find Null-Polarization Models with Artifacts

## Corpus Analysis to Determine Undecidable Replications

To analyze published replications we need to expand our notation to index individual *experiments*
with replications, and from which different journal articles, track and analyze it,

## Evaluating Grouop Polarization Experimental Designs