### 2025-11-25 GroupPolarizationStatmod/paper/method.qmd
started on specify sig thresh

---
title: Method
editor: 
  markdown: 
    wrap: 80
author: Matt Turner · [https://mat.phd](https://mat.phd)
date: today
toc: true
bibliography: this.bib
---

\newcommand{\mupre}{\mu_\mathrm{pre}}
\newcommand{\mupost}{\mu_\mathrm{post}}
\newcommand{\sigmapre}{\sigma_\mathrm{pre}}
\newcommand{\sigmapost}{\sigma_\mathrm{post}}

\newcommand{\thetafit}{\theta^\mathrm{fit}}

\newcommand{\fdr}{\mathrm{FDR}}

\newcommand{\meanobs}{\bar{o}}
\newcommand{\meanobsemp}{\bar{\mathbf{o}}}
\newcommand{\meanobst}{\meanobs_t}
\newcommand{\meanobsjt}{\meanobs_{j,t}}
\newcommand{\meanobspre}{\meanobs_\mathrm{pre}}
\newcommand{\meanobspost}{\meanobs_\mathrm{post}}
\newcommand{\tpre}{t_\mathrm{pre}}
\newcommand{\tpost}{t_\mathrm{post}}

A group polarization replication is proven undecidable, and therefore not truly
a replication, if we can find a model of *mere agreement* that appears to be
polarized under that replication's experimental design. If we cannot find such a
model, we cannot determine a replication to be undecidable. Mere agreement in
this model is a dynamic where agreement increases within the group, but
extremism does not—in other words, A *null-polarization distributions* is a set
of two Gaussian distributions: one representing the pre-discussion opinions of
the group, and the other representing the post-discussion opinions. Our
*null-polarization distributions* represent the process of the group finding
*mere agreement*, which is when the average group opinion is constant over time,
but opinion variance decreases to represent agreement (or, equivalently,
*consensus* [@DeGroot1974]).

# Latent Opinion Model

Throughout this paper we model a participant $i$'s latent opinion at time $t$ as
a normally-distributed random variable,
$$ 
\omega_{i,t} \sim
\mathcal{N}(\mu_t, \sigma_t),
$$ {#eq-random-variable-def}
with mean $\mu_t$ and standard deviation $\sigma_t$. 
Group polarization in terms of latent opinions is then defined as any pair of pre- and
post-discussion opinion distributions where the mean increases in magnitude over
time—formally, $| \mupost | > | \mupre |$. We need to take magnitudes since radicalization may be towards increasingly negative or positive opinion values.

For group polarization experiments, only two times are relevant:
*pre-discussion*, $t=\mathrm{pre}$, and *post-discussion*, $t=\mathrm{post}$. Since participant groups are assumed to be likeminded, we could also assume they come to agree more over time as well as become more extreme in our group polarization model, or in terms of the latent opinion variance, $\sigmapre > \sigmapost$—but only one study in our ten actually measured opinion variance to check this [@Schkade2010].

A *null-polarization model* is a model that is consistent with the *negation*, or *null*, of the  *group polarization hypothesis*. The group polarization hypothesis is that group discussion causes the average group opinion to increase, assuming the group is aligned on a particular issue, though opinion strength varies. Therefore a null model is one where the latent mean stays
constant, i.e., $\mu = \mu_\mathrm{pre} = \mu_\mathrm{post}$.

*Mere agreement* is the type of null-polarization model we use here. Mere agreement is a null-model because the mean latent opinion is set to be constant before and after discussion. We represent *agreement*—or equivalently, *consensus*—as a decrease in the standard deviation that defines the pre- and post-discussion latent opinion distributions. The mere agreement model is the cornerstone of our analysis, since it is empirically plausible and generates *spurious* observations of group polarization under mere agreement models, due to measurement artifacts induced by clipping all latent opinion values outside the measurement range—we call this *spurious polarization* for short. We simulate opinion measurements in two different ways, but both use the mere agreement model of latent opinions to test whether and by how much the measurement procedure distorts null-polarization to look like group polarization.

::: {style="text-align: center;"}
\[ TABLE \[tab:latent-variables\](#tab:latent-variables){reference-type="ref"
reference="tab:latent-variables"} ABOUT HERE \]
:::

## Two Ways to Model Measurement that Extend the Latent Opinion Model

We simulate measurement in two different ways—i.e., with two different models—for the two parts of our analysis. The first
is the *expectation model*: an
efficient way to *simulate measurement of the observed average* opinion of a group, using the *mathematical expectation value* of the distribution of measured opinions as a proxy. In the second
model, the *stochastic model*, we simulate individual participant responses by first drawing an opinion for each participant from the latent opinion distribution (@eq-. The opinions are binned to the appropriate ordinal value.

We use the expectation model for its
efficiency in the first step of our analysis: to find a *null-polarization
model* that also fits the data for each experimental condition studied—if we
find one that means we must retract this replication since this proves that the
group polarization is undecidable with the data we have. Once we find any such
undecidable results, we use the null-polarization model to help us further
evaluate the experimental design itself.

In the *stochastic model*, we simulate group polarization data for the
*Evaluation* step of the analysis. We simulate experimental data, with the same
number of participants as the original experiment, using the null-polarization
model found in the *Identification* step, if one was found.

We *evaluate* the experimental design, $e$, in terms of the false positive rate,
also known as the Type I error rate or family-wise error rate, denoted
$\alpha_e$. We close our analysis by demonstrating how our method can be
inverted to *evaluate* experimental designs in the pre-registration stage to
avoid preventable replication failures like we identify here.

The measurement models are introduced in detail along with the corresponding stage of the analysis that uses them. First, the *Identification* stage that uses the *expectation* measurement model; second, the *Evaluation* stage that uses the *stochastic* measurement model.

# *Identification* of Invalid Replications

In the first stage of our analysis, the *Identification* stage, we check if we can *invalidate* a replication of group polarization achieved in some experiment/experimental condition, which we label $e$. Our invalidation strategy is to find a mere agreement null-polarization model, call it $\nu$, that predicts the exact observations reported in each experimental replication in our corpus. If one is found, then we have found a model consistent with the *negation* of the group polarization hypothesis that explains the data as well as the group polarization hypothesis itself. In other words, the hypothesis is not proved or disproved given the data, but *undecidable*: the data cannot *decide* the hypothesis one way or another because of potential measurement artifacts. Because the experiment failed to produce a valid replication, the results are effectively stripped of evidentiary value and warrant exclusion from the aggregate evidence base.

But how do we find a $\nu$ that generates spurious group polarization? And how do we even simulate the group polarization experiments and measurements that make mere agreement in latent opinions look like group polarization? We simulate 
experimental measurements of group polarization by using the mathematical *expectation value* of latent opinions defined by $\nu$ as a proxy for the observed average value of ordinal-valued reported opinions (described in the first subsection below). We then use this measurement model in a numerical root-finding algorithm that identifies a $\nu$ that matches published observations, if possible. If we find such a $\nu$, this means we have a model
consistent with the null hypothesis that explains published observations. In other words, the replication is invalidated since the group polarization hypothesis is *undecidable* given the data—i.e., it cannot decide whether the hypothesis or its null is correct.

So, to review, this subsection introduces how the *Identification* stage of the analysis works. In this stage we identify for which experiments, $e$, does a mere agreement null model look like group polarization under $e$'s measurement procedure.

## The Expectation Model of Opinion Measurement

We use the *expectation model* of opinion measurement to *Identify* which, if any, mere agreement null-polarization models generate spurious group polarization in a particular experiment, $e$. The trick to this model, and the source of its name, is that we use the theoretical *expected value* of opinion measurements as a proxy for the *observed average* of participant opinions under an experimental design, $e$.

We simulate the measurement of the mean observed opinion at time $t$ 
by first calculating integrating the normal probability density function,  $f(\omega)$, over measurement bin boundaries, over each bin's corresponding range of opinion values. For example, consider an ordinal scale  that runs from -3 (strongly disagree) to +3 (strongly agree) (as
used by @Moscovici1969, for example). the frequency with which the "2" bin is
chosen is just the total area underneath the probability density curve between
1.5 and 2.5.

All bins have a width of 1 in latent opinion space, except for the two most
extreme bins at opposite ends of the sentiment spectrum---this works because bin
values are always separated by 1. All continuous latent opinion values
$\omega < b_1 + \frac{1}{2} = \theta_1$ get mapped to the ordinal measurement
value $b_1$---so, $b_1$'s lower threshold is $\theta_0 = -\infty$. Similarly,
all opinions $\omega > b_B - 0.5 = \theta_B$ are reported as the maximum bin
value, $b_B$, with an effective upper bound of $\theta_{B+1} = +\infty$. We can
define all thresholds compactly as [@Liddell2018] 
$$
\theta_j = \begin{cases}
  -\infty         & \text{ if } j=0 \\
  \infty          & \text{ if } j=B \\
  b_j + \frac{1}{2} & \text{ otherwise.}
\end{cases}
$$

The probability density that a participant reports the opinion in bin index $i$ at time $t$ is
$$
f_{it}(\mu, \sigma_t) = \int_{\theta_{i - 1}}^{\theta_{i}} f(\omega; \mu, \sigma_t)~d\omega,
$$
which we calculated in R to evaluate the integral at the two threshold values $\theta_{i-1}$ and $\theta_i$. Note that the normal probability density function we used is parameterized by mean $\mu$ and standard deviation $\sigma_t$.
$$
f(\omega; \mu, \sigma_t) = 
    \frac{1}{\sigma_t \sqrt{2 \pi}} e^{-\frac{(\omega - \mu)^2}{2 \sigma_t^2}}.
$$

To simulate the measurement of the *average* observed opinion at time $t$, $\bar{o}_t$, as the expected value of the probability density function of observed ordinal opinions. In other words, we simulate $\bar{o}_t$ by calculating the weighted sum of products between the probability of observing each bin value and the bin value itself:
$$
\bar{o}_t(\nu, e) = \sum_{i=1}^B b_i f_{it}(\nu, e).
$${#eq-sim-ave-observed}

Group polarization is the difference between the post-deliberation mean observed
opinion and the pre-deliberation one. We will calculate the magnitude of spurious group polarization under a mere agreement null model represented by its collected distribution parameters, $\nu = (\mu, \sigmapre, \sigmapost)$, under experimental design, $e$: 
$$
g(\nu, e) = |\bar{o}_\mathrm{post}(\nu, e)| - |\bar{o}_\mathrm{pre}(\nu, e)|
$${#eq-spurious-gp}

In the *Inspect* stage of our analysis we will determine which group polarization replications are invalidated due to equifinality between the *mere agreement null-polarization hypothesis* on the one hand, and the *group polarization hypothesis*
on the other. This requires more than just showing there is some $\nu$ that causes spurious polarization for some experiment where $g(\nu, e) > 0$. We have to find a $\nu$ that generates the exact $g(\nu, e)$ reported in the original journal article *and* check that simulated $\bar{o}_\mathrm{pre}$ and $\bar{o}_\mathrm{pre}$ match published values.

::: {text-align="center"}
\[ TABLE
[\[tab:experiment-parameters\]](#tab:experiment-parameters){reference-type="ref"
reference="tab:experiment-parameters"} ABOUT HERE \]
:::

::: center
\[ TABLE
[\[tab:simulated-observations\]](#tab:simulated-observations){reference-type="ref"
reference="tab:simulated-observations"} ABOUT HERE \]
:::

## Example: the "Americans" condition in Moscovici and Zavalloni (1969)

One of the replications we invalidate here is from Moscovici and Zavalloni (1969),
the "Americans" condition as they called it—we coded it `Moscovici1969_Americans`. 
They report that participant opinions polarized after discussing the survey topic, "American economic aid is always used for political pressure." They report survey 
responses increased in extremity from an average of -0.9 before discussion to
-1.6 after, on a seven-bin scale
from -3 indicating "strongly disagree" to +3 indicating "strongly agree".

Moscovici and Zavalloni say their results confirm the basic hypothesis, that \emph{group discussion causes polarization}. But we found a mere agreement null-polarization model that generates spurious group polarization using our model—specifically, $\nu = (\mu = -1.2, \sigma_\mathrm{pre}= 4.3, and \sigma_\mathrm{post}= 1.9)$. This means that the hypothesis is actually undecidable given the data, invalidating Moscovici and Zavalloni's replication.
When a hypothesis and its null explain data equally well, the data tell us nothing.

Now let's use the expectation model to prove that this $\nu$ generates spurious group polarization in this experiment, i.e., $g(\nu, e= \mathrm{Moscovici1969\_Americans}) > 0$. 
In the process we demonstrate how the expectation model works to simulate the measurement of spurious group polarization (@eq-spurious-gp).

First, we calculate the probability density vector using the [`makeProbVec`](https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/model.R#L44) function in the file [`GroupPolarizationStatmod/model.R`](https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/model.R) in the [project code repository on GitHub](https://github.com/mt-digital/gp-statmod) for both experiment times: pre- and post-discussion:
$$
f_\mathrm{pre} = 
\begin{pmatrix}
  0.38 & 0.09 & 0.09 & 0.09 & 0.08 & 0.07 & 0.19
\end{pmatrix}
$$ 
and
$$
f_\mathrm{post} = \
\begin{pmatrix} 
0.25 & 0.19 & 0.21 & 0.17 & 0.11 & 0.05 & 0.03 
\end{pmatrix}.
$$

When we plug in the bin values and copmuted $f_\mathrm{pre}$ values into the simulated average pre-discussion opinion definition (@eq-sim-ave-observed),
$$
\bar{o}_\mathrm{pre} = -3 \cdot 0.38~~+-2\cdot 0.09~~+\ldots+3\cdot0.19 = -0.6
$$
$$
\bar{o}_\mathrm{post} = -3 \cdot 0.25~~+-2\cdot 0.19~~+\ldots+3\cdot0.03 = -1.04
$$
So, we find that for the `Moscovici1969_Americans`, our null-polarization model appears polarized, with the exact same average pre- and post-discussion opinions reported by Moscovici and Zavalloni (1969), and the same magnitude of radicalization (CHECK THE PRESENTATION OF ALL THREE ABOVE)
$$
g(\nu, e) = \bar{o}_\mathrm{post} - \bar{o}_\mathrm{pre} = -1.04 - (-0.6) = - 0.4.
$$

::: {style="text-align: center;"}
FIGURE [\[fig:distros\]](#fig:distros){reference-type="ref"
reference="fig:distros"} ABOUT HERE 
:::

Since the simulated average opinions match the observed ones—note that as a consequence, the spurious polarization magnitude also matches the observed one.
In other words, we *identified* a null model that explains the data. 
If both a hypothesis and its null explain data equally well, then the hypothesis is *undecidable*

In this example we started with

## Root-finding Algorithm to Find Null-Polarization Models with Artifacts

# Evaluation of Experimental Designs

- The evaluation stage complements our *Identify* stage by filling in more details about the experimental design.

## False positive rate

First, how frequently do the $\nu$ that we found generate simulated data
  that fool a rigorous statistical design that properly accounts for the measurement process? This serves two roles: 
  1. Confirm that the $\nu$ we found could create simulated datasets that appear polarized under statistical analysis, which is different from using the expected value of an exact latent opinion distribution, integrated with the bin values to simulate the average opinion.
  2. Evaluate the underlying experimental design independent of the failure to account for how ordinal measurements inevitably clip extreme latent opinions.
  3. We do this by simulating experiment data using $\nu$ we identified in the previous stage. We simulate 1000 experiments indexed by $i$ for each experiment $e$.

To calculate the false detection rate we begin by simulating 
pre- and post-deliberation 
opinions by drawing $N_e$ samples from normal latent opinion
distributions that generated spurious group polarization, i.e., one sample at each experiment stage for each simulated participant. The latent opinion distributions are specified with the mere agreement null model labelled $\nu_e$.

Each trial $i$ of $N_e$ simulated pre- and post-deliberation observations is then inspected to see
if spurious group polarization occurred in the simulation. Spurious group
polarization is determined by first inferring the latent parameters of each
simulated experimental trial by fitting an ordered probit model~\cite{Liddell2018}. The procedure infers the latent distribution parameters that we in fact specified with $\nu$: $\mu_\mathrm{pre}$, $\sigma_\mathrm{pre}$, $\mu_\mathrm{post}$, and $\sigma_\mathrm{post}$.

The inferred parameters are then used to calculate the effect size for that
trial, Cohen's $d$, defined as~\cite[p. 331]{Liddell2018}
$$
  d = \frac{\mupost - \mupre}{\sqrt{\frac{1}{2}(\sigmapost^2 + \sigmapre^2)}}.
$$ {#eq-cohens}
It is implied that these values are the inferred values, not the values from $\nu$, when used in the context of calculating Cohen's $d$. Note $d=0$ always for the true $\nu$.

The family-wise error rate is the fraction of trials 
with $d_{ei} > d^*$, written
$$
  \alpha_e = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
$$ {#eq-study-aggregate}
where
$$
  \Theta(d_{ei}; d^*) = 
    \begin{cases}
      1 & \text{ if } d_{ei} > d^* \\
      0 & \text{ otherwise.}
    \end{cases}
$$
$d^*$ is the *significance threshold*, which must be specified by the experimenter.  There are three common significance levels which we adopt here: $d^* = 0.2$ for "low", $d^*=0.5$ for "medium", and $d^* = 0.8$ for "high".

To calculate the family-wise error rate
across all experimental conditions of a study, $s$, 
\begin{equation}
  \alpha_s = \frac{1}{|s|} \sum_{e \in s} \alpha_e,
\end{equation}
\noindent
where $e \in s$ indicates sum over all experimental conditions $e$ presented in the article associated with study $s$.
The mean across all results identified as plausibly spurious is 
\begin{equation}
  \alpha_\mathrm{all} = \frac{1}{|E|} \sum_{e \in E} \alpha_e.
\end{equation}

In the final part of our *Evaluation* analysis we adapt this analysis to re-align the question of statistical rigor: we can be specific now, and instead of wondering, "what significance level is truly rigorous", we can use our techniques to make the choice of $d^*$ to conform to a specific design requirement. Specifically we will calculate what $d^*$ must be in order to limit $\alpha \leq 0.05$ as an example.

## False discovery rate

We finally may calculate the false discovery rate by plugging this $\alpha$
into Equation~\ref{eq:fdr} and assuming empirically-motivated values for 
base rate, $b$, and statistical power, $W$, as well as more pessimistic and
optimistic settings for base rate and power.

Using the notation outlined above, the false detection rate is
\begin{equation}
  \fdr = \frac{(1 - b)\alpha}{(1 - b)\alpha + bW}.
  \label{eq:fdr}
\end{equation}

## *Specify* Significance Threshold for Low False Positive Rates

In the final analysis step we calculate what significance threshold, $d^*$,
would limit the false positive rate to 0.05 or less.

There are many possible
null models we could choose.

what significance threshold would have limited inverting this procedure to find what significance value
$d^*$ is necessary to achieve a low family-wise error rate $\alpha_e = 0.05$ for
experimental condition, $e$.

To do this we first calculate $\alpha_e(d^*)$ for a
sufficiently wide range of $d^*$. We then find the lowest $d^*$ for which
$\alpha_e(d^*) \leq 0.05$.