### 2025-11-25 GroupPolarizationStatmod/paper/method.qmd
magic expanding methods section

---
title: Method
editor: 
  markdown: 
    wrap: 80
author: Matt Turner · [https://mat.phd](https://mat.phd)
date: today
toc: true
bibliography: this.bib
format:
  html:
    grid:
      body-width: 880px  
---

\newcommand{\mupre}{\mu_\mathrm{pre}}
\newcommand{\mupost}{\mu_\mathrm{post}}
\newcommand{\sigmapre}{\sigma_\mathrm{pre}}
\newcommand{\sigmapost}{\sigma_\mathrm{post}}

\newcommand{\thetafit}{\theta^\mathrm{fit}}

\newcommand{\fdr}{\mathrm{FDR}}

\newcommand{\meanobs}{\bar{o}}
\newcommand{\meanobsemp}{\bar{\mathbf{o}}}
\newcommand{\meanobst}{\meanobs_t}
\newcommand{\meanobsjt}{\meanobs_{j,t}}
\newcommand{\meanobspre}{\meanobs_\mathrm{pre}}
\newcommand{\meanobspost}{\meanobs_\mathrm{post}}
\newcommand{\tpre}{t_\mathrm{pre}}
\newcommand{\tpost}{t_\mathrm{post}}

A group polarization replication is proven undecidable, and therefore not truly
a replication, if we can find a model of *mere agreement* that appears to be
polarized under that replication's experimental design. If we cannot find such a
model, we cannot determine a replication to be undecidable. Mere agreement in
this model is a dynamic where agreement increases within the group, but
extremism does not—in other words, A *null-polarization distributions* is a set
of two Gaussian distributions: one representing the pre-discussion opinions of
the group, and the other representing the post-discussion opinions. Our
*null-polarization distributions* represent the process of the group finding
*mere agreement*, which is when the average group opinion is constant over time,
but opinion variance decreases to represent agreement (or, equivalently,
*consensus* [@DeGroot1974]).

# Latent Opinion Model

Throughout this paper we model a participant $i$'s latent opinion at time $t$ as
a normally-distributed random variable, $$ 
\omega_{i,t} \sim
\mathcal{N}(\mu_t, \sigma_t),
$$ {#eq-random-variable-def} with mean $\mu_t$ and standard deviation
$\sigma_t$. Group polarization in terms of latent opinions is then defined as
any pair of pre- and post-discussion opinion distributions where the mean
increases in magnitude over time—formally,
$| \mu_\mathrm{post}| > | \mu_\mathrm{pre}|$. We need to take magnitudes since
radicalization may be towards increasingly negative or positive opinion values.

For group polarization experiments, only two times are relevant:
*pre-discussion*, $t=\mathrm{pre}$, and *post-discussion*, $t=\mathrm{post}$.
Since participant groups are assumed to be likeminded, we could also assume they
come to agree more over time as well as become more extreme in our group
polarization model, or in terms of the latent opinion variance,
$\sigma_\mathrm{pre}> \sigma_\mathrm{post}$—but only one study in our ten
actually measured opinion variance to check this [@Schkade2010].

A *null-polarization model* is a model that is consistent with the *negation*,
or *null*, of the *group polarization hypothesis*. The group polarization
hypothesis is that group discussion causes the average group opinion to
increase, assuming the group is aligned on a particular issue, though opinion
strength varies. Therefore a null model is one where the latent mean stays
constant, i.e., $\mu = \mu_\mathrm{pre} = \mu_\mathrm{post}$.

*Mere agreement* is the type of null-polarization model we use here. Mere
agreement is a null-model because the mean latent opinion is set to be constant
before and after discussion. We represent *agreement*—or equivalently,
*consensus*—as a decrease in the standard deviation that defines the pre- and
post-discussion latent opinion distributions. The mere agreement model is the
cornerstone of our analysis, since it is empirically plausible and generates
*spurious* observations of group polarization under mere agreement models, due
to measurement artifacts induced by clipping all latent opinion values outside
the measurement range—we call this *spurious polarization* for short. We
simulate opinion measurements in two different ways, but both use the mere
agreement model of latent opinions to test whether and by how much the
measurement procedure distorts null-polarization to look like group
polarization.

::: {style="text-align: center;"}
\[ TABLE \[tab:latent-variables\](#tab:latent-variables){reference-type="ref"
reference="tab:latent-variables"} ABOUT HERE \]
:::

# Two Models of Group Polarization Measurement

We simulate measurement in two different ways—i.e., with two different
models—for different parts of our analysis. The first is the *expectation
model*: an efficient way to *simulate measurement of the observed average*
opinion of a group, using the *mathematical expectation value* of the
distribution of measured opinions as a proxy. In the second model, the
*stochastic model*, we simulate individual participant responses by first
drawing an opinion for each participant from the latent opinion distribution
(@eq-random-variable-def). The opinions are binned to the appropriate ordinal
value.

We use the expectation model for its efficiency in the first step of our
analysis: to find a *null-polarization model* that also fits the data for each
experimental condition studied—if we find one that means we must retract this
replication since this proves that the group polarization is undecidable with
the data we have. Once we find any such undecidable results, we use the
null-polarization model to help us further evaluate the experimental design
itself.

In the *stochastic model*, we simulate group polarization data for the
*Evaluation* step of the analysis. We simulate experimental data, with the same
number of participants as the original experiment, using the null-polarization
model found in the *Identification* step, if one was found.

We *evaluate* the experimental design, $e$, in terms of the false positive rate,
also known as the Type I error rate or family-wise error rate, denoted
$\alpha_e$. We close our analysis by demonstrating how our method can be
inverted to *evaluate* experimental designs in the pre-registration stage to
avoid preventable replication failures like we identify here.

The measurement models are introduced in detail along with the corresponding
stage of the analysis that uses them. First, the *Identification* stage that
uses the *expectation* measurement model; second, the *Evaluation* stage that
uses the *stochastic* measurement model.

# Part 1: *Identify* Invalid Replications

In the first stage of our analysis, we try to *identify* a null polarization
models, $\nu$, that invalidate published detections of group polarization if the
measurement procedure induces *spurious group polarization* for that null
polarization model. Null polarization can take many forms—all that is required
is that the latent mean opinion stays constant before and after discussion. Here
we use a *mere agreement* null polarization model, where we simulate
*consensus*, the tendency for sufficiently similar opinions to become more
similar—but we leave out polarization. In mere agreement, moderate and extreme
opinions approach change by equal magnitude towards each other. We therefore
represent consensus as a decrease in the standard deviation of the latent
opinion distribution from pre- to post-discussion. We simulate the measurement
of the group's average opinion before and after discussion and measure the
spurious group polarization, if there is any. We use this procedure in a search
algorithm we developed to identify which combination of constant latent mean and
pre- and post-discussion opinion variances generate spurious group polarization,
if there are any.

Our invalidation strategy, then, is to find a mere agreement null polarization
model, $\nu$, that predicts the exact observations reported in each experimental
replication in our corpus. If one is found, then we have found a model
consistent with the *negation* of the group polarization hypothesis that
explains the data as well as the group polarization hypothesis itself. In other
words, the hypothesis is not proved or disproved given the data, but
*undecidable*: the data cannot *decide* the hypothesis one way or another
because of potential measurement artifacts. Because the experiment failed to
produce a valid replication, the results are effectively stripped of evidentiary
value and warrant exclusion from the aggregate evidence base. We apply our
analysis to every experimental condition across all ten journal articles in our
corpus, which we label $e$.

But how do we find a $\nu$ that generates spurious group polarization? And how
do we even simulate the group polarization experiments and measurements that
make mere agreement in latent opinions look like group polarization? We simulate
experimental measurements of group polarization by using the mathematical
*expectation value* of latent opinions defined by $\nu$ as a proxy for the
observed average value of ordinal-valued reported opinions (described in the
first subsection below). We then use this measurement model in a numerical
root-finding algorithm that identifies a $\nu$ that matches published
observations, if possible. If we find such a $\nu$, this means we have a model
consistent with the null hypothesis that explains published observations. In
other words, the replication is invalidated since the group polarization
hypothesis is *undecidable* given the data—i.e., it cannot decide whether the
hypothesis or its null is correct.

So, to review, this subsection introduces how we identify which experimental
data can be explained by a mere agreement null model, where the ordinal survey
measurement procedure induces a clipping artifact that hides changes in extreme
opinions that fall outside the measurement scale.

## The Expectation Model of Spurious Group Polarization

We use the *expectation model* of opinion measurement to *Identify* which, if
any, mere agreement null-polarization models generate spurious group
polarization in a particular experiment, $e$. The trick to this model, and the
source of its name, is that we use the theoretical *expected value* of opinion
measurements as a proxy for the *observed average* of participant opinions under
an experimental design, $e$.

We simulate the measurement of the mean observed opinion at time $t$ by first
calculating integrating the normal probability density function, $f(\omega)$,
over measurement bin boundaries, over each bin's corresponding range of opinion
values. For example, consider an ordinal scale that runs from -3 (strongly
disagree) to +3 (strongly agree) (as used by @Moscovici1969, for example). the
frequency with which the "2" bin is chosen is just the total area underneath the
probability density curve between 1.5 and 2.5.

All bins have a width of 1 in latent opinion space, except for the two most
extreme bins at opposite ends of the sentiment spectrum---this works because bin
values are always separated by 1. All continuous latent opinion values
$\omega < b_1 + \frac{1}{2} = \theta_1$ get mapped to the ordinal measurement
value $b_1$---so, $b_1$'s lower threshold is $\theta_0 = -\infty$. Similarly,
all opinions $\omega > b_B - 0.5 = \theta_B$ are reported as the maximum bin
value, $b_B$, with an effective upper bound of $\theta_{B+1} = +\infty$. We can
define all thresholds compactly as [@Liddell2018] $$
\theta_j = \begin{cases}
  -\infty         & \text{ if } j=0 \\
  \infty          & \text{ if } j=B \\
  b_j + \frac{1}{2} & \text{ otherwise.}
\end{cases}
$$

The probability density that a participant reports the opinion in bin index $i$
at time $t$ is $$
f_{it}(\mu, \sigma_t) = \int_{\theta_{i - 1}}^{\theta_{i}} f(\omega; \mu, \sigma_t)~d\omega,
$$ which we calculated in R to evaluate the integral at the two threshold values
$\theta_{i-1}$ and $\theta_i$. Note that the normal probability density function
we used is parameterized by mean $\mu$ and standard deviation $\sigma_t$. $$
f(\omega; \mu, \sigma_t) = 
    \frac{1}{\sigma_t \sqrt{2 \pi}} e^{-\frac{(\omega - \mu)^2}{2 \sigma_t^2}}.
$$

To simulate the measurement of the *average* observed opinion at time $t$,
$\bar{o}_t$, as the expected value of the probability density function of
observed ordinal opinions. In other words, we simulate $\bar{o}_t$ by
calculating the weighted sum of products between the probability of observing
each bin value and the bin value itself: $$
\bar{o}_t(\nu, e) = \sum_{i=1}^B b_i f_{it}(\nu, e).
$$ {#eq-sim-ave-observed}

Group polarization is the difference between the post-deliberation mean observed
opinion and the pre-deliberation one. We will calculate the magnitude of
spurious group polarization under a mere agreement null model represented by its
collected distribution parameters,
$\nu = (\mu, \sigma_\mathrm{pre}, \sigma_\mathrm{post})$, under experimental
design, $e$: $$
g(\nu, e) = |\bar{o}_\mathrm{post}(\nu, e)| - |\bar{o}_\mathrm{pre}(\nu, e)|
$$ {#eq-spurious-gp}

In the *Inspect* stage of our analysis we will determine which group
polarization replications are invalidated due to equifinality between the *mere
agreement null-polarization hypothesis* on the one hand, and the *group
polarization hypothesis* on the other. This requires more than just showing
there is some $\nu$ that causes spurious polarization for some experiment where
$g(\nu, e) > 0$. We have to find a $\nu$ that generates the exact $g(\nu, e)$
reported in the original journal article *and* check that simulated
$\bar{o}_\mathrm{pre}$ and $\bar{o}_\mathrm{pre}$ match published values.

::: {text-align="center"}
\[ TABLE
[\[tab:experiment-parameters\]](#tab:experiment-parameters){reference-type="ref"
reference="tab:experiment-parameters"} ABOUT HERE \]
:::

::: center
\[ TABLE
[\[tab:simulated-observations\]](#tab:simulated-observations){reference-type="ref"
reference="tab:simulated-observations"} ABOUT HERE \]
:::

## Example: Invalidating the "Americans" condition in Moscovici and Zavalloni (1969)

One of the replications we invalidate here is from Moscovici and Zavalloni
(1969), the "Americans" condition as they called it—we coded it
`Moscovici1969_Americans`. They report that participant opinions polarized after
discussing the survey topic, "American economic aid is always used for political
pressure." They report survey responses increased in extremity from an average
of -0.9 before discussion to -1.6 after, on a seven-bin scale from -3 indicating
"strongly disagree" to +3 indicating "strongly agree".

Moscovici and Zavalloni say their results confirm the basic hypothesis, that
\emph{group discussion causes polarization}. But we found a mere agreement
null-polarization model that generates spurious group polarization using our
model—specifically,
$\nu = (\mu = -1.2, \sigma_\mathrm{pre}= 4.3, and \sigma_\mathrm{post}= 1.9)$.
This means that the hypothesis is actually undecidable given the data,
invalidating Moscovici and Zavalloni's replication. When a hypothesis and its
null explain data equally well, the data tell us nothing.

Now let's use the expectation model to prove that this $\nu$ generates spurious
group polarization in this experiment, i.e.,
$g(\nu, e= \mathrm{Moscovici1969\_Americans}) > 0$. In the process we
demonstrate how the expectation model works to simulate the measurement of
spurious group polarization (@eq-spurious-gp).

First, we calculate the probability density vector using the
[`makeProbVec`](https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/model.R#L44)
function in the file
[`GroupPolarizationStatmod/model.R`](https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/model.R)
in the [project code repository on
GitHub](https://github.com/mt-digital/gp-statmod) for both experiment times:
pre- and post-discussion: $$
f_\mathrm{pre} = 
\begin{pmatrix}
  0.38 & 0.09 & 0.09 & 0.09 & 0.08 & 0.07 & 0.19
\end{pmatrix}
$$ and $$
f_\mathrm{post} = \
\begin{pmatrix} 
0.25 & 0.19 & 0.21 & 0.17 & 0.11 & 0.05 & 0.03 
\end{pmatrix}.
$$

When we plug in the bin values and copmuted $f_\mathrm{pre}$ values into the
simulated average pre-discussion opinion definition (@eq-sim-ave-observed), $$
\bar{o}_\mathrm{pre} = -3 \cdot 0.38~~+-2\cdot 0.09~~+\ldots+3\cdot0.19 = -0.6
$$ $$
\bar{o}_\mathrm{post} = -3 \cdot 0.25~~+-2\cdot 0.19~~+\ldots+3\cdot0.03 = -1.04
$$ So, we find that for the `Moscovici1969_Americans`, our null-polarization
model appears polarized, with the exact same average pre- and post-discussion
opinions reported by Moscovici and Zavalloni (1969), and the same magnitude of
radicalization (CHECK THE PRESENTATION OF ALL THREE ABOVE) $$
g(\nu, e) = \bar{o}_\mathrm{post} - \bar{o}_\mathrm{pre} = -1.04 - (-0.6) = - 0.4.
$$

::: {style="text-align: center;"}
FIGURE [\[fig:distros\]](#fig:distros){reference-type="ref"
reference="fig:distros"} ABOUT HERE
:::

Since the simulated average opinions match the observed ones—note that as a
consequence, the spurious polarization magnitude also matches the observed one.
In other words, we *identified* a null model that explains the data. If both a
hypothesis and its null explain data equally well, then the hypothesis is
*undecidable* under the available data. This invalidates this experiment's claim
to be a replication of support for the group polarization hypothesis. We
replicate this analysis for all 57 experimental conditions that purport to
support the group polarization hypothesis. To facilitate this analysis, we
created a web app to manage the data entry, the process of identifying invalid
replications, and managing the outcome data of which replications were
invalidated, and what null model parameters generated the spurious polarization.

## System for identifying invalid replications

The next step is to scale up the analysis we did manually in the example above,
including develop a systematic way to find, and not guess, the set of null model
parameters that

![**Computational system to interactively simulate opinion change, identify
invalid replications, and evaluate experimental
designs.**](Figures/Model/SoftwareMap_Mashup.png){#fig-computational-system
fig-align="center" width="1011"}

# Part 2: *Evaluate* Experimental Designs

In the previous step we identified which, if any, null polarization models
generate spurious group polarization with the expectation model of opinion
measurement. This is enough to invalidate a replication, but it leaves
unanswered questions. First, we would like to know how what the chance was that
a published detection of group polarization was actually a false positive,
assuming that participant opinions were distributed according to the null model
identified in the first part of the analysis. This is equivalent to the false
positive rate of the experimental design, which we denote $\alpha_e$ for
experiment $e$. Second, we would like to estimate the false discovery rate for
our corpus, which is a good proxy for the field as a whole, since we selected
top quality, highly influential work that has been cited thousands of times. The
false discovery rate in this case tells us the probability that a published
detection of group polarization is actually a false detection. (@eq-fdr).

## Generative Model of Spurious Group Polarization

To evaluate an experimental design's ability to properly identify null effects,
we need to simulate participant responses assuming the mere agreement null model
that invalidated the replication, if one was found. We simulate data from a
single group polarization experiment by sampling participants' ordinal opinions
at pre-/post-discussion from the pre- and post-discussion opinion distributions
specified by the null model that invalidated the replication in that experiment,
$\nu_e$. If there are $N_e$ participants in experiment $e$ then our simulated
data sets contained $2N_e$ rows. We recorded $N_e$ in the web app as metadata
along with data about average opinions and opinion shifts for each experimental
condition across the corpus.

It can be helpful to

## False positive rate

To calculate the false detection rate we begin by simulating pre- and
post-deliberation opinions by drawing $N_e$ samples from normal latent opinion
distributions that generated spurious group polarization, i.e., one sample at
each experiment stage for each simulated participant. The latent opinion
distributions are specified with the mere agreement null model labelled $\nu_e$.

Each trial $i$ of $N_e$ simulated pre- and post-deliberation observations is
then inspected to see if spurious group polarization occurred in the simulation.
Spurious group polarization is determined by first inferring the latent
parameters of each simulated experimental trial by fitting an ordered probit
model [@Liddell2018]. The procedure infers the latent distribution parameters
that we in fact specified with $\nu$: $\mu_\mathrm{pre}$, $\sigma_\mathrm{pre}$,
$\mu_\mathrm{post}$, and $\sigma_\mathrm{post}$.

The inferred parameters are then used to calculate the effect size for that
trial, Cohen's $d$, defined as\~\cite[p. 331]{Liddell2018} $$
  d = \frac{\mu_\mathrm{post}- \mu_\mathrm{pre}}{\sqrt{\frac{1}{2}(\sigma_\mathrm{post}^2 + \sigma_\mathrm{pre}^2)}}.
$$ {#eq-cohens} It is implied that these values are the inferred values, not the
values from $\nu$, when used in the context of calculating Cohen's $d$. Note
$d=0$ always for the true $\nu$.

The family-wise error rate is the fraction of trials with $d_{ei} > d^*$,
written $$
  \alpha_e = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
$$ {#eq-study-aggregate} where $$
  \Theta(d_{ei}; d^*) = 
    \begin{cases}
      1 & \text{ if } d_{ei} > d^* \\
      0 & \text{ otherwise.}
    \end{cases}
$$ $d^*$ is the *significance threshold*, which must be specified by the
experimenter. There are three common significance levels which we adopt here:
$d^* = 0.2$ for "low", $d^*=0.5$ for "medium", and $d^* = 0.8$ for "high".

To calculate the family-wise error rate across all experimental conditions of a
study, $s$, $$
\alpha_s = \frac{1}{|s|} \sum_{e \in s} \alpha_e,
$$ where $e \in s$ indicates sum over all experimental conditions $e$ presented
in the article associated with study $s$. The mean across all results identified
as plausibly spurious is $$
 \alpha_\mathrm{all} = \frac{1}{|E|} \sum_{e \in E} \alpha_e.
$$

In the final part of our *Evaluation* analysis we adapt this analysis to
re-align the question of statistical rigor: we can be specific now, and instead
of wondering, "what significance level is truly rigorous", we can use our
techniques to make the choice of $d^*$ to conform to a specific design
requirement. Specifically we will calculate what $d^*$ must be in order to limit
$\alpha \leq 0.05$ as an example.

## False discovery rate

The false discovery rate is the false positive rate, scaled by the rate of true
negative effects and normalized by the frequency of true positive effects\~
(@eq-fdr) and assuming empirically-motivated values for base rate, $b$, and
statistical power, $W$, as well as more pessimistic and optimistic settings for
base rate and power. Using this notation, the false detection rate is

$$
\mathrm{FDR}= \frac{(1 - b)\alpha}{(1 - b)\alpha + bW}.
$$ {#eq-fdr}

# Part 3: *Specify* Significance Threshold for Experimental Design

In the final analysis step we calculate what significance threshold, $d^*$,
would limit the false positive rate to 0.05 or less.

There are many possible null models we could choose.

what significance threshold would have limited inverting this procedure to find
what significance value $d^*$ is necessary to achieve a low family-wise error
rate $\alpha_e = 0.05$ for experimental condition, $e$.

To do this we first calculate $\alpha_e(d^*)$ for a sufficiently wide range of
$d^*$. We then find the lowest $d^*$ for which $\alpha_e(d^*) \leq 0.05$.