### 2025-11-26 GroupPolarizationStatmod/paper/Discussion.qmd
updated method, started Discussion standalone

---
title: Discussion
editor: 
  markdown: 
    wrap: 80
author: Matt Turner · [https://mat.phd](https://mat.phd)
date: today
toc: true
bibliography: this.bib
format:
  html:
    grid:
      body-width: 880px  
---

\newcommand{\mupre}{\mu_\mathrm{pre}}
\newcommand{\mupost}{\mu_\mathrm{post}}

\newcommand{\sigmapre}{\sigma_\mathrm{pre}}
\newcommand{\sigmapost}{\sigma_\mathrm{post}}

\newcommand{\thetafit}{\theta^\mathrm{fit}}

\newcommand{\fdr}{\mathrm{FDR}}

\newcommand{\meanobs}{\bar{o}}
\newcommand{\meanobsemp}{\bar{\mathbf{o}}}
\newcommand{\meanobst}{\meanobs_t}
\newcommand{\meanobsjt}{\meanobs_{j,t}}
\newcommand{\meanobspre}{\meanobs_\mathrm{pre}}
\newcommand{\meanobspost}{\meanobs_\mathrm{post}}
\newcommand{\tpre}{t_\mathrm{pre}}
\newcommand{\tpost}{t_\mathrm{post}}

# The Full Bayesian Rigor: Deriving Effect Size Posteriors and Justifying the FPR Approximation

In evaluating the False Positive Rate (FPR) of a simulated null model, the primary challenge is moving beyond the raw parameter posterior distributions ($p(\mu_1 | \text{data})$, $p(\sigma_2 | \text{data})$, etc.) to accurately quantify the posterior distribution of the derived effect size, $d$. This section outlines the statistically rigorous method for this derivation and justifies the pragmatic approximation necessary for the evaluation stage of this paper.

## The Correct Derivation of $\boldsymbol{\delta}$

In our model, the effect size $d$ is a function of four estimated parameters ($\mu_1, \mu_2, \sigma_1, \sigma_2$), defined by the authors whose work we replicate as:

$$d = \frac{\mu_1 - \mu_2}{\sqrt{(\sigma_1^2 + \sigma_2^2)/2}}$$

The statistically rigorous method for obtaining the posterior distribution of $d$, denoted $p(d | \text{data})$, is through **marginalization by sampling** across the joint posterior distribution of the primary parameters $p(\mu_1, \mu_2, \sigma_1, \sigma_2 | \text{data})$.

### Necessity of MCMC Samples

The parameters are fundamentally **sequentially dependent** in the MCMC chain, as the estimation of means is conditional on the estimation of variances. This dependency is captured in the joint posterior distribution.

To maintain statistical rigor, we must avoid calculating $d$ by combining the isolated summary statistics (e.g., the posterior median of $\mu_1$ and the posterior median of $\sigma_1$) because this procedure, known as **plug-in estimation**, ignores the critical correlation between parameters, resulting in an unquantified and likely underestimated uncertainty.

The correct procedure requires using the full MCMC samples:

1.  For each usable iteration $k$ of the Markov Chain (where $K$ is the total number of samples), a plausible set of values is drawn: $(\mu_1^{(k)}, \mu_2^{(k)}, \sigma_1^{(k)}, \sigma_2^{(k)})$.
2.  The effect size is calculated for that specific set of values:
    $$d^{(k)} = \frac{\mu_1^{(k)} - \mu_2^{(k)}}{\sqrt{\frac{(\sigma_1^{(k)})^2 + (\sigma_2^{(k)})^2}{2}}}$$
3.  The set $\{d^{(1)}, d^{(2)}, \ldots, d^{(K)}\}$ then forms the complete posterior distribution $p(d | \text{data})$.

The center of this distribution, the **posterior median ($\tilde{d}$)**, is the most robust point estimate for the effect size, as it minimizes expected absolute error and is less sensitive to potential skewness than the posterior mean.

## Justifying the FPR Approximation

Due to the computational complexity and the unavailability of the original MCMC chains for all $57,000$ fits, a pragmatic simplification was necessary for the subsequent evaluation of the False Positive Rate (FPR).

### The FPR Criterion

A standard method for estimating the FPR (or Type I error rate) in simulation studies is to count the proportion of null trials where the observed effect size meets or exceeds a predefined threshold, $d^*$.

$$\text{FPR} = \frac{\text{Number of Null Fits where } |d| > d^*_{\text{threshold}}}{\text{Total Number of Null Simulations}}$$

In this analysis, $d^*_{\text{threshold}}$ is set to the commonly adopted small effect size threshold of $d^*=0.2$ (and varied for sensitivity analysis), avoiding the reliance on approximated $95\%$ credible intervals, which would require further, less reliable approximations.

### The Compromise: Plug-in and Heuristics

We employed a simplified plug-in value for $d$ using the stored posterior means:

$$d_{\text{plug}} = \frac{\hat{\mu}_1 - \hat{\mu}_2}{\sqrt{(\hat{\sigma}_1^2 + \hat{\sigma}_2^2)/2}}$$

Where $\hat{\theta}$ represents the posterior mean of the parameter $\theta$.

While this approach is a statistical shortcut, the error is mitigated by two factors: the use of the $d^*_{\text{threshold}}$ heuristic, and the large simulated sample size ($N=500$ per group), which ensures that the posteriors for $\mu$ and $\sigma$ are sufficiently narrow. This minimizes the difference between $d_{\text{plug}}$ and the true posterior median $\tilde{d}$, making the shortcut defensible as a pragmatic necessity given the study's scale.