### 2025-12-04 GroupPolarizationStatmod/paper/ms.tex
cleaned up and simplified for faster build in texshop

% ==============================================================================
% CLEANED ms.tex (Optimized for XeLaTeX/Fira Fonts)
% - Consolidated preamble and ms.tex.
% - Removed all conflicting packages (lmodern, redundant font settings).
% - Uses Fira Sans/Mono/Math consistently.
% ==============================================================================

\documentclass[preview]{article}

% --- CORE FONT SETUP (Sans-Serif Headings, Serif Body, Professional Math) ---
\usepackage{fontspec}        
\usepackage{unicode-math}    
%
%% 1. Main Text: Set to Latin Modern Roman (Stable, Professional Serif)
%\setmainfont{Latin Modern Roman} 
%
% 2. Monospaced/Code Text: Fira Mono
%\setmonofont{Fira Mono} 
%
%% 3. Mathematical Notation: Latin Modern Math (Stable, Conventional Serif Math)
%\setmathfont{Latin Modern Math}

% 4. Heading Font Setup (Fira Sans)
\newfontfamily{\sansheadings}{Fira Sans} 
\usepackage{sectsty}
\sectionfont{\sansheadings \raggedright \Large}
\subsectionfont{\sansheadings \raggedright \large}
\subsubsectionfont{\sansheadings \raggedright \normalsize}
% Ensure your title command uses \sansheadings as defined in the previous step.
% ...

% --- ESSENTIAL PACKAGES (Cleaned and Prioritized) ---

% 2. Core Text Utilities and Geometry
\usepackage[margin=1.25in,letterpaper]{geometry} % Defines page layout
\usepackage[]{microtype}    % High-quality font protrusion and spacing
\usepackage{upquote}        % Straight quotes in verbatim/code
\usepackage{calc}           % Utility for calculations

% 3. Graphics, Color, and Table Fundamentals
\usepackage[table,dvipsnames,svgnames,x11names]{xcolor} % Color definitions
\usepackage{graphicx}       % For including graphics
\usepackage{booktabs}       % Professional quality tables
\usepackage{longtable}      % Multi-page tables
\usepackage{array}          % Enhanced array and column definitions
\usepackage{multirow}       % Merging rows in tables
\usepackage{makecell}       % Formatting cells in tables

% 4. Caption Handling (Load basic caption definitions BEFORE floatrow)
\usepackage{caption}        
\usepackage{subcaption}     
\usepackage{floatrow}       % Complex float/caption placement and formatting

% 5. Math Environments and Algorithm Structure
\usepackage{amsmath}        % Essential math environments (align, gather, etc.)
\usepackage{algorithm}
\usepackage{algpseudocode}

% 6. Hyperlinks and References
\usepackage{hyperref}       % Hyperlinks and PDF metadata (must be loaded relatively late)
\usepackage{bookmark}       % PDF bookmark control
\usepackage{xurl}           % Better URL line breaks
\usepackage{footnotehyper}  % Footnotes integration with hyperref

% 7. Custom Utilities (Needed for your document structure)
\usepackage{iftex}          
\usepackage{authblk}        % Author affiliations
\usepackage{etoolbox}       % Utility commands for patching/macros
\usepackage{wrapfig}        % Text wrapping around figures

% --- End of ESSENTIAL PACKAGES ---

% --- BIBLIOGRAPHY CONFIGURATION ---
% Use the biblatex settings you had in preamble.tex
\usepackage[sorting=none, style=numeric-comp]{biblatex}
\addbibresource{/Users/mt/workspace/Writing/library.bib}
%\addbibresource{this.bib}

% --- FLOATROW / TABLE SETUP (From ms.tex) ---
\DeclareFloatFont{small}{\small}
\floatsetup[table]{font=small}

% --- HYPERREF SETUP (From ms.tex) ---
\hypersetup{
  pdftitle={Group polarization replications may be marred by high false discovery rates},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue}}

% --- CUSTOM COMMANDS (From ms.tex) ---
\newcommand{\mupre}{\mu_\mathrm{pre}}
\newcommand{\mupost}{\mu_\mathrm{post}}
\newcommand{\sigmapre}{\sigma_\mathrm{pre}}
\newcommand{\sigmapost}{\sigma_\mathrm{post}}
\newcommand{\thetafit}{\theta^\mathrm{fit}}
\newcommand{\fdr}{\mathrm{FDR}}
\newcommand{\meanobs}{\bar{o}}
\newcommand{\meanobsemp}{\bar{\mathbf{o}}}
\newcommand{\meanobst}{\meanobs_t}
\newcommand{\meanobsjt}{\meanobs_{j,t}}
\newcommand{\meanobspre}{\meanobs_\mathrm{pre}}
\newcommand{\meanobspost}{\meanobs_\mathrm{post}}
\newcommand{\tpre}{t_\mathrm{pre}}
\newcommand{\tpost}{t_\mathrm{yopost}}
\definecolor{myorange}{RGB}{240, 96, 0}
\newcommand{\mt}[1]{{\textcolor{myorange} {({\tiny MT:} #1)}}}

% --- TITLE AND AUTHOR INFO (From ms.tex) ---
\makeatletter
\title{\Large If the Null Fits, You Must Omit: Measurement Artifacts Undermine Group Polarization Replications}
\makeatother
\author[1,*]{{Matthew A.~Turner}}
\affil[1]{\small Environmental Social Sciences, Stanford Doerr School of Sustainability, Stanford University}
\author[2,3]{{Paul E.~Smaldino}}
\affil[2]{\small Cognitive and Information Sciences, University of California, Merced}
\affil[3]{\small Santa Fe Institute} 
\affil[*]{\small Correspondence: \href{mailto:maturner@stanford.edu}{maturner@stanford.edu}}
\date{\today}

% ==============================================================================
% BEGIN DOCUMENT
% ==============================================================================
\begin{document}
\maketitle

\renewcommand*\contentsname{\small Table of contents}
{
\hypersetup{linkcolor=gray}
\setcounter{tocdepth}{3}
\footnotesize
\tableofcontents
}

% Center rule to separate TOC from rest of paper
\vspace{-2em}
\begin{center} \noindent\rule{4cm}{0.4pt} \end{center}
\vspace{-2em}

% --- INPUT CHAPTERS ---
\input{intro.tex}
\input{method.tex}
\input{analysis.tex}
\input{discussion.tex}
\input{tables.tex}
\input{figures.tex}

\clearpage

\printbibliography[title=References]

\appendix

\renewcommand{\thefigure}{A\arabic{figure}}
\setcounter{figure}{0}
\clearpage

%%%%%%%%%% Supplement %%%%%%%%%%
\pagebreak
\begin{center}
  \textbf{\Large \textsf{Group polarization replications may be marred by high false discovery rates (Supplementary Material)}}
\end{center}

%%%%%%%%%% Prefix a "S" to all equations, figures, tables and reset the counter %%%%%%%%%%
\setcounter{equation}{1}
\setcounter{figure}{0}
\setcounter{section}{0}
\setcounter{table}{0}
\setcounter{page}{1}
\makeatletter
\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\thepage}{S\arabic{page}}
\makeatother

\section{Analysis}

\begin{figure}
  \caption{Boxplot of Cohen's $d$ for the ordered probit model over 1000
    simulation trials for each experimental condition (y-axis in A and B).
    Each trial represents a possible outcome of a group polarization experiment where
    the true opinion shift is zero. The mean across trials is closer to 0 than
    in the metric case, but many simulated zero-shift experiments still result in
    an observed shift. This illustrates the weak statistical power of these group
    polarization experiments.}
  \label{fig:OrdinalBoxplot}
  \centering
    \includegraphics[width=1.0\textwidth]{Figures/Analysis/ordinal_cohens.pdf}
\end{figure}

\end{document}