### 2025-12-11 index.qmd
updated docs

---
title: Notes and code examples to accompany "If the Null Fits, You Must Omit"
author: Matt Turner · [Email Me!](mailto:matt@mat.phd) · [https://mat.phd](https://mat.phd) 
date: today
toc: true
toc-expand: 2
editor: 
  markdown: 
    wrap: 90
bibliography: "this.bib"

execute:
  # echo: false     # hide code
  warning: false  # hide warnings
  message: false  # hide messages
  # error: false
---

<!-- Define the tex command for the null model -->
\newcommand{\nullmodel}{\mathcal{M}_0}

<!-- Begin main document -->
# Introduction

This document supports my article with Paul Smaldino, *Null Fits? Must Omit!* It follows the structure of the article's _Method_ and _Analysis_ sections, providing a tutorial explanation, first, of the code used to model group polarization experiments and opinion measurement; and, second, how the code works to summarize, analyze, and visualize simulation results.

Our article describes how we analyzed simulations of group polarization experiments to show that ~95% of experimental replications of the group polarization effect are invalid due to non-identifiability from measurement artifacts across ten journal articles that comprised our curated corpus. There were a total of 57 independent experimental conditions across these studies.

We furter estimated at least a 70% chance that any published replication of group polarization is, in fact, a false detection.

We also show how this simulation system can be used constructively to specify what effect size should be used as a detection threshold to limit FPR to 0.05. For more details of the motivation, theory, and results of our analysis, please see our paper.

**You can read this documentation online, hosted by GitHub:** https://mt-digital.github.io/gp-statmod

**Or, you can read this document and run the code examples yourself.** To follow along and run code on your own, please clone the `gp-statmod` directory:

```{bash}
#| eval: false
git clone https://github.com/mt-digital/gp-statmod.git
```

Change directories to go to the `gp-statmod` directory where this document lives:

```{sh}
#| eval: false
cd gp-statmod
```

Then you can run the examples in this document, edit them, etc., on your own computer.

Either way, you will learn how the code works that simulates measurements of opinion change
in group polarization experiments that test if and by how much group discussion polarizes opinions. I used this modeling system to simulate 57 different group polarization experimental conditions across a corpus of ten journal articles published over several decades, and across social science subdisciplines for my paper with Paul Smaldino, "Null Fits? You Must Omit! How Measurement Artifacts Undermine Many Group Polarization Replications".

# Method

I developed two models of experimental observations in group polarization experiments
to serve complementary purposes. These models provide two ways to simulate observations of _mere agreement_, which serves as our null-polarization model, $\nullmodel$,

First is the _expectation_ model where we model the observation of the average group opinion as the _expectation value_ of the distribution of ordinal opinions. We calculate the ordinal distribution of opinions by writing the endpoints of each measurement bins as limits of integration in a series of $B$ integrals, where $B$ is the number of measurement bins. For example, an example of an ordinal measurement scale might be a Likert-style series of disagree-agree levels, associated with numerical values:
-3 indicates

## Expectation model of group polarization experiments

See Method section of the paper where we explain how to calculate the probability density
vector representing the probability that a participant gives the opinion represented by
each measurement bin, $b$.

Here we show how to calculate the mean observed opinion at pre- and post-discussion to
show how spurious group polarization occurs under a null polarization hypothesis.

@Moscovici1969 report observing a mean opinion of -0.6 in pre-discussion opinions and mean
opinion -1.04 post-discussion, measured using an ordinal scale from -3 (strongly disagree)
to +3 (strongly agree). Let's calculate the probability density vector for the pre- and
post-discussion distributions then calculate the mean opinion for each one. The latent
distributions have the same mean, -1.2, but different variances, which results in a
shifted observed mean due to opinion clipping.

![**Mean group opinions pre- and post-discussion from Moscovici & Zavalloni
(1969).**](images/clipboard-1443885641.png){width="575"}

We call this a *null polarization model* (defined in terms of latent opinions) that
generates *spurious group polarization.* *Spurious group polarization* is when the
observed opinion distributions appear to polarize (i.e., *radicalize*, or *become more
extreme*) that does notrepresent the true change in latent opinions. The measurement
procedure *clips* extreme opinions less than the minimum bin value, or greater than the
maximum, causing the average observed opinion to *spuriously* appear polarized.

Let's use some modeling code from the repository to demonstrate. For this part we need to
source (*use code from*) the `model.R` file that has the the
[`binProb`](https://github.com/mt-digital/gp-statmod/blob/final-draft/GroupPolarizationStatmod/model.R#L7)
function I use below. Also make sure you have set the working directory to the
`GroupPolarizationStatmod` directory within the `gp-stats` project/repository directory.

For a preliminary step, adapt my code below for your computer. Change the argument to
`root.dir` to be the location of the `GroupPolarizationStatmod` directory on your own
computer.

Now in the following chunk, the `source` function will know where to look for the file
with the `makeProbVec` function we need, `model.R`.

```{r warning=FALSE}
# This is how I set the working directory
# Adapt for your own computer
# setwd("~/workspace/gp-statmod/GroupPolarizationStatmod/")
# Load makeProbVec from file model.R
source("model.R")
```

Now read in the pre- and post-discussion means from `Moscovici1969_Americans`

```{r}
# Experimental means
mosc1969_ave_pre <- -0.6
mosc1969_ave_post <- -1.04
```

Create a hypothesized null model that will spuriously look like polarization, with the
same pre- and post-discussion means above. These correspond to $\mu$,
$\sigma_\mathrm{pre}$, and $\sigma_\mathrm{post}$ from the paper that define a null
polarization model, $\nullmodel = (\mu, \sigma_\mathrm{pre}, \sigma_\mathrm{post})$.

```{r}
# Hypothesized null model that looks polarized
mu <- -1.2
sd_pre <- 4.3
sd_post <- 1.9
```

The function `makeProbVec` creates the vector $f[b]$ of probabilities that bin $b_i \in b$
will be reported by a randomly-selected participant, given the measurement bin values and
the distribution of latent opinions, assumed normal, defined by the mean and standard
deviation, $\mu$ and $\sigma$. and bin values. Here bin values are $b=-3,-2,\ldots,3$;
$\mu=-1.2$; and $\sigma=4.3$.

```{r}
# Vector w/ probability density for each bin
f_pre <- makeProbVec(-3:3, -1.2, 4.3)
# Print to two significant digits
sprintf("%.2f", f_pre)
```

Let's `barplot` to have a look:

```{r out.width=500}
barplot(f_pre, 
        names.arg = -3:3,
        ylim = c(0, 0.4),
        ylab = "Pct. Responding",
        xlab = "Bin Value")
```

Let's check the simulated average observed opinion, $\bar{o}$, using the expected value of
the distribution, i.e., we assume $\bar{o} = \langle o \rangle$, so we use the standard
formula for expected value, which assumes that $f[b]$ is normalized.

$$
\bar{o} = \sum_{b=b_1}^{b_B} b \cdot f[b].
$$

In code,

```{r}
# Calculate simulated and compare with actual
sim_obs_pre <- sum((-3:3) * f_pre)
print(abs(sim_obs_pre - mosc1969_ave_pre))
```

This is within one significant digit greater than the reporting scale, meaning it is an
acceptable fit.

This shows that even just a little spread in the distribution of latent opinions results
in a seriously distorted mean [@Liddell2018]—one that reproduces the mean reported by
@Moscovici1969, but actually confirming the null polarization hypothesis!

Let's check the simulated average observed opinion for post-deliberation, repeating the
previous steps for $t=\mathrm{post}$.

```{r}
f_post <- makeProbVec(-3:3, mu, sd_post)
sprintf("%.2f", f_post)
# Calculate simulated and compare with actual
sim_obs_post <- sum((-3:3) * f_post)
print(abs(sim_obs_post - mosc1969_ave_post))
```

Here we get even closer agreement between the observed post-discussion average opinion in
@Moscovici1969 and the one we simulated.

# The Generative Model and Bayesian Inference for False Detections

The analysis of the generative model has two major components:

1.  **Calculate false positive/detection rates for the simulated null polarization
    model**, *mere agreement*—where polarization experiment participant opinions agree
    more following discussion than before. Recall this is modeled as a decrease in the
    standard deviation of latent opinions, but a constant mean latent opinion. Here, then, we calculate the false positive rate, $\alpha_e$ as a function of the effect-detection threshold, $d^*$, i.e., 
    $$
      \alpha_e(d^*) = \Pr\!\left( \hat{d} \ge d^{*} \mid \mathcal{M}_{0} \right).
    $$ {#eq-false-positive-rate}
    In the study, we calculate the right-hand side as the frequency of the 1000 trials where $\hat{d} \ge d^*$ for simulations of experiment, $e$. 
2.  **Find an effect–detection threshold**, $d^*(\alpha)$, such that, under the null
    model, the probability of observing an estimated effect size exceeding
    \$d\^\*(\\alpha)\$ is some target false positive rate, $\alpha^* = 0.05$, for example: 
    $$
     d^* = \min~d,~~\text{subject to}~\Pr\!\left( \hat{d} \ge d \mid \mathcal{M}_{0} \right)  \le\alpha^*.
     $$ {#eq-threshold-spec}
     This yields a null-model–calibrated false–positive rate for detecting a group
    polarization effect.

The first step in our _generative_ model of false detections is to _generate_ simulated 
experimental data from a mere agreement null-polarization model. Then we can calculate how frequently an effect size was inferred that exceeded the effect-detection threshold.

## Simulate Experimental Data

I simulate experimental group polarization data for a given time, $t$, with latent group opinion mean and standard deviation, $\mu_t$ and $\sigma_t$, and with a survey response scale defined by $B$ bin values, $b_i$, with $i=1,\ldots,B$.

<!-------- CONTINUE HERE 12/11! ------>

## Simulate each experimental design's false positive rate

Here's how to calculate the Type I or family-wise error rate for all the experiments
(though I misnamed the function back in the day, so it calls all of the error rates
"family-wise" because I had understood them to be synonyms).

```{r}
source("src/fwer.R")

probit_fits_dir <- "data/probit_fits"
# Set a "moderate" significance, according to Cohen's definitions 
significance_val <- 0.5
fwer_tbl <- calculate_fwer(probit_fits_dir, significance_val);

print(head(fwer_tbl, n = 10))
```

This is used inside the `calc_fdr_vs_significance` function that adds another column to
this dataframe, the false discovery rate (FDR) for a set of significance values. Here it
is being called with the default metascience parameters with a base rate of $b=0.1$ and a
statistical power of $W=0.8$. We consider three effect-detection thresholds in the paper: "small", "medium", and "large" effect sizes traditionally assumed practically
meaningful across social sciences [@CohenBook1988].

The `calc_fdr_vs_significance` function calculates the false-positive rate ($\alpha$) and false detection rate ($\mathrm{FDR}$) reported for a variety of effect-detection thresholds, $d^*$, as we call them in the paper. When I first wrote the code, though, I was thinking about detection thresholds as synonymous with "significance values". So that's why the function name, `calc_fdr_vs_significance`, includes the word `significance`, which is a stand-in for the maximally descriptive "effect-detection thresholds" terminology. This seemed fine to keep things simple and not change the name everywhere in the code. But it was right to change it for the paper, since *significance* is only understood in a p-value based  *null-hypothesis significance testing* approach, where a hypothesis is confirmed if the p-value is less than some particular threshold. I updated the prose in the article to reflect this distinction, but left the code as-is.

I put this function in the `scripts` directory because to run the full analyses presented in the paper it's only necessary to run `calc_fdr_vs_significance` since it creates a column for the false positive rate and the estimated false detection rate calculated for each experimental design we studied.

Below I call the function with the defaults specified, for illustration. Remember the `significance_vals` map to _effect-detection thresholds_ ($d^*$) in the paper.

```{r}
source("scripts/analysis.R")

# This 
fdr_over_thresholds <- calc_fdr_vs_significance(
  probit_fits_dir = probit_fits_dir,  # defined in above code block
  base_rate = 0.1, power = 0.8, significance_vals = c(0.2, 0.5, 0.8)
)

print(head(fdr_over_thresholds, n=10))
```
To see `SigVal = 0.8` (ie $d^* = 0.8$), check the `tail` of the table:
```{r}
print(tail(fdr_over_thresholds, n = 10))
```

## Find effect-detection thresholds to achieve $\alpha = 0.05$

In the second step of our generative model analysis I calculate the solution to
the problem in @eq-threshold-spec, finding the effect-detection threshold, $d^*$, that achieves a target false positive rate, $\alpha^* = 0.05$. We need an updated dataset of Type I error rates calculated with a finer-grained set of candidate effect-detection thresholds, which we just call $d$ if they're candidates, than the simple three, 0.2, 0.5, and 0.8, used in the previous section.

```{r}
# The fwer is equivalently Type I error—written before I understood the difference,
# not worth changing the name for this long-running project before it's done.
det_thresh_tbl <- sigval_for_low_fwer(fdr_over_thresholds, target_fwer = 0.05)

print(det_thresh_tbl)
```

# Analysis

## _Identify_ Invalid Replications: All But 5% Were Invalidated

## _Evaluate_ Experimental Designs: ~70% Chance a Polarization Detection was Erroneous

## _Specify_ Effect-Detection Thresholds: Spec Experiments—Don't Just Design Them—and Avoid Informed Guesswork

# References